---
title: Dynamic Documention for "The Effects of a Minimum-Wage Increase on Employment
  and Family Income"
author: "Download/Modify/Contribute the analysis [here](https://github.com/fhoces/dd_mw)"
date: '`r paste("Last edit:", Sys.Date(), sep=" ")`'
output:
  html_document:
    code_folding: show
    collapsed: no
    keep_md: yes
    number_sections: yes
    smooth_scroll: no
    theme: united
    toc: yes
    toc_depth: 2
    toc_float: yes
editor_options: 
  chunk_output_type: console
---

<!-- 
The following chunk is to hide/display code with a specific title ('R' and 'Stata')
-->
  <script language="javascript"> 
    function toggle(num) {
      var ele = document.getElementById("toggleText" + num);
      var text = document.getElementById("displayText" + num);
      if(ele.style.display == "block") {
        ele.style.display = "none";
        text.innerHTML = "show";
      }
      else {
        ele.style.display = "block";
        text.innerHTML = "hide";
      }
   } 
  </script>

<!-- 
Following line is to get rpubs to recognize external links-->
<base target="_top"/>
<!-- 
resource_files:
- apps/app1/app.R
runtime: shiny
-->


# Introduction   

The role of policy analysis is to connect research with policy. Because of heavy time constrains, policy analyses are typically ambiguous regarding the details of how the analysis was carried out. This creates three problems: (i) its hard to understand the connection between research and policy, (ii) allows policy makers to cherry pick policy reports, and (iii) hinders systematic improvement and/or automation of parts of the analysis. In this document we demonstrate the use of a reproducible workflow to reduce the ambiguity in policy analysis.

Here we attempt to contribute to the policy discussion of the minimum wage. The minimum wage is a contentious policy issue in the US. Increasing it has positive and negative effects that different policymakers value differently. We aim to add clarity on what those effects are, how much do we know about them, and how those effects vary when elements of the analysis change. We select the most up-to-date, non-partisan, policy analysis of the effects of raising the minimum wage, and build an open-source reproducible analysis on top of it.

In 2014 the Congressional Budget Office published the report titled ["The Effects of a Minimum-Wage Increase on Employment and Family Income"](https://www.cbo.gov/publication/44995). The report receive wide attention from key stakeholders and has been used extensible as an input in the debate around the minimum wage. To this date we consider the CBO report to be the best non-partisan estimation of the effects of raising the minimum wage at the federal level. Although there was disagreement among experts around some technical issues, this disagreement has been mainly circumscribed around one of the many inputs used in the analysis, and we can fit the opposing positions in to our framework.

Our purposes are twofold: First, promote the technical discussion around a recurrent policy issue (minimum wage) by making explicit and visible all the components and key assumptions of its most up-to-date official policy analysis. Second, demonstrate how new scientific practices of transparency and reproducibility (T & R) can be applied to policy analysis. We encourage the reader to collaborate in this document and help develop an ever-improving version of the important policy estimates[^3] (re)produced here.

To achieve our goal we reviewed the CBO report and extract the key components of its analysis. We adapt new guidelines propose by the scientific community  ([TOP](https://cos.io/top/)) into policy analysis. In it, the analysis achieves the highest standards of transparency and reproducibility (T & R) when the data, methods and workflow are completely reproducible and every part of the analysis and its assumptions, are easily readable. We also benefit from hindsight and structure this document around the costs and benefits mainly discussed in the policy debate. 

CBO's report, in its original form already represents a significant improvement in T & R relative to the standard practices of policy analyses. The report contains most of the components required for a full reproduction. We add the missing components, make explicit assumptions when needed, complement the narrative explanations with some mathematical formulae, visualizations, and the analytical code use behind all the replication.

**Important Note:**

Although our aim is to translate practices of T & R from Science to Policy Analysis, we need to highlight an important difference regarding reproducibility between the two of them. A scientific report takes the form of a peer review publication that represent several months or years of research, followed up by a review process that can be as lengthy as the research itself. For this reason, when a scientific publication is subject to replication is expected to succeed. Policy analysis is usually performed under tight deadlines, and is not unusual to rely on arbitrary assumptions and/or irreproducible calculations. For this reasons we do not attempt to replicate the CBO report as a way of testing the veracity of the analysis. We use reproducibility, paired with full transparency, to generate a living document that represents the best policy analysis up to date. Our expectations are that this living document will be serve as a building block to discuss and incorporate incremental improvements to the policy analysis used to inform the debate around the minimum wage. 

The CBO report describes three policy estimates: the effects of raising the minimum wage on income of families with members that receive a raise, the effects on income of families with members that loose their jobs, and the distributions of losses in the economy used to pay for the raise in the minimum wage. All the policy estimates to replicate are presented in the following tables.

**Note on the code languages (`R` and `Stata`):** The analysis can be replicated using either language, but only R provides the one-click workflow. For `Stata` the reader has to copy and paste the scripts sequentially or excecute [this do file](link to final do file). 


```{r setup, include=FALSE}
dir.create(file.path(tempdir(), ".checkpoint"), recursive = TRUE, showWarnings = FALSE)

#library(checkpoint)
#checkpoint("2018-07-31", checkpointLocation = tempdir())


# Loading required libraries
list.of.packages <- c("foreign", "weights", "survey", "Hmisc", 
                      "openxlsx", "rio", "highr", "XML", "RCurl", "treemap", 
                      "reshape2", "tidyverse", "xtable", "haven", "here", "rvest")

new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages, repos= "http://cran.cnr.berkeley.edu/") 

lapply(list.of.packages, require, character.only = TRUE)
# Setting working directory
library(here)

# Setting up workflow
for (folder in c("rawdata", "data", "documentation", "output", "paper", "scripts")) {
  if ( !(folder %in% dir()) ) {
    dir.create(folder)    
  }
}

## NEED TO ADD CHECKPOINT

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache = FALSE)

statapath <- "/Applications/Stata/Stata.app/Contents/MacOS/Stata"

display_code <- TRUE #set false when outputing to pdf (maybe?)
```


```{r table with output,  eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}  
# This is summary of all the policy estimates presented in the report
# Aggregate effects
output.template1 <- matrix(" ???",nrow = 6, ncol = 1)
rownames(output.template1) <- c("wage gains (billions of $)", "wage losses (bns of $)", "Balance losses (bns of $)", 
                                "Net effect (bns of $)", "# of Wage gainers (millions)", "#of Wage losers (millions)")
colnames(output.template1) <- "Effects/Policy Estimates"

output.template1[,"Effects/Policy Estimates" ] <- c("31", "~5", "~24" , "2", "16.5", "0.5")

#Some distributional effects
output.template2 <- matrix(" ???",nrow = 2, ncol = 4)
rownames(output.template2) <- c("Balance losses (bns of $)", "Net effect (bns of $)")
colnames(output.template2) <- c("<1PL", "[1PL, 3PL)", "[3PL, 6PL)", ">6PL")

output.template2["Balance losses (bns of $)", ] <- c("~0.3", "~3.4", "~3.4", "~17")
output.template2["Net effect (bns of $)", ] <- c("5", "12", "2", "-17")

knitr::kable(
  list(
    output.template1  ),
  caption = 'Policy estimates in CBO report: Overall effects', booktabs = TRUE, 
  align = 'c'
)

knitr::kable(
  list(
    output.template2
  ),
  caption = 'Policy estimates in CBO report: Distributional effects across poverty lines (PL)', booktabs = TRUE, 
  align = 'c'
)





#This is the first suggestion: change how policy estimates are reported and add information.
mod.output.template <- matrix(" ???",nrow = 7, ncol = 5)
rownames(mod.output.template) <- c("wage gains", "wage losses", "Balance losses", "Net effect", "# of Wage gainers", "#of Wage losers", "Population")
colnames(mod.output.template) <- c("<1PL", "[1PL, 3PL)", "[3PL, 6PL)", ">6PL", "Total")

mod.output.template[,"Total" ] <- c("31", "~5", "~24" , "2", "16.5", "0.5", "330/140")
mod.output.template["Balance losses", ] <- c("~0.3", "~3.4", "~3.4", "~17", "~24")
mod.output.template["Net effect", ] <- c("5", "12", "2", "-17", "2")


#knitr::kable(mod.output.template, caption="Template for final results to replicate", digits = 1)
```



```{r SA params,  eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}  
# The base growth rate of wages will determine the dispersion of wage growth, the mean wage groth rate will be given by the 10 year economic forecast. 
#data inputs:
param.wage.gr <- 1.0
param.worker.gr <- 1 

param.eta.lit <- 1.0 #
param.factor.extrap <- 1.0
param.base.growth <- 0.024 * 1.0
param.N <- 1.0
param.fract.minwage <- 1.0
param.noncomp <- 1.0
param.F.adj <- 1.0
param.av.wage.var <- 1.0
param.wages <- 1.0 ####
param.nonwage.gr <- 1.0
param.hours <- 1.0
param.weeks <- 1.0
param.factor.1 <- 1
param.net.benef <- 2e9*1.0
param.ripple <- c("scope_below" = 8.7*1, "scope_above" = 11.5*1.0, "intensity" = 0.5*1.0)
param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.2, 0.4, 0.40) #c(0.39567218, 0.53851506, 0.06581275) 
param.jobcut <- 1
param.states.raise <- 1.0


```


In this companion we attempt to reproduce all the policy estimates of table 1 and 2, and walk the reader through all the details behind it. At a high level, two key components need to be computed: employment effects and distributional effects. 


# Employment effects
At a general level the effects on employment ($\widehat{\Delta E}$) will be calculated using a more detailed version of the following equation:

$$
\begin{equation}
\widehat{\Delta E} = N \times \eta \times \% \Delta w  + \text{Other factors}
\label{eq:1}
\tag{1}
\end{equation}
$$
  
Where $N$ represents the relevant population, $\eta$ the elasticity of labor demand, $\Delta w$ the relevant percentual variation in wages, and the *Other factors* will encapsulate effects on employment through an increase in the aggregate demand.  

We first describe some general methodological considerations and then describe how to compute each component from equation $\eqref{eq:1}$.  

## Data, wages, and forecast  
We describe the data used, the chosen wage variable, and the procedure used to forecast the wage and population distribution of 2016 using data from 2013.  

### Data
The Current Population Survey (CPS) was used to compute the effects on employment. From the analysis described in the section on distributional effects of the original report, we can deduce that the data corresponds to the Outgoing Rotation Group (ORG). CPS is a monthly cross sectional survey. The same individual is interviewed eight times over a period of 12 months. The interviews take place in the first and last 4 months of that period. By the 4th and 12th interview, individuals are asked detailed information on earnings. The CPS ORG file contains the information on this interviews for a given year. We analyze the data for 2013.   

Currently three versions of these data sets can be found online: [CPS raw files](http://thedataweb.rm.census.gov/ftp/cps_ftp.html#cpsbasic), [ORG NBER](http://www.nber.org/data/morg.html) and [ORG CEPR](http://ceprdata.org/cps-uniform-data-extracts/cps-outgoing-rotation-group/). The analysis will be performed using the CPER ORG data base.

The weights used in our analysis will be `orgwgt/12`

#### Code to load the data 

<a id="displayText" href="javascript:toggle(1);">`R`</a>  
<div id="toggleText1" style="display: none">  
  
```{r loading cps data R, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE}
call.cps.org.data <- function(){
  data_use <- "CPER_ORG"
  # Using CEPR ORG data 
  if (data_use == "CPER_ORG") {
  # Checking if working directory contains data, download if not. 
    if ( !("cepr_org_2013.dta" %in% dir(path = "./rawdata/") ) ) {
    	# create name of file to store data
    	tf <- "cepr_org_2013.zip"
    
    	# download the CPS repwgts zipped file to the local computer
    	download.file(url =  "http://ceprdata.org/wp-content/cps/data/cepr_org_2013.zip", tf , mode = "wb" )
    
    	# unzip the file's contents and store the file name within the temporary directory
    	fn <- unzip(exdir = "./rawdata/", zipfile = tf , overwrite = T )
    }
    df <- haven::read_dta("./rawdata/cepr_org_2013.dta")
  }

  # Using NBER ORG data 
  if (data_use == "NBER_ORG") {
    # Checking if working directory contains data, download if not. 
    if ( !("morg13.dta" %in% dir()) ) {
      # Downloading data 53mb
      df <- haven::read_dta("http://www.nber.org/morg/annual/morg13.dta")
    }
    df <- haven::read_dta("morg13.dta")
  }
  
  df <- tbl_df(df)
  
  # There are 1293 cases with missin values for the weigths. I delete them from the data. 
  df <- df %>% filter(!is.na(orgwgt))
  df$final_weights <- df$orgwgt/12
  df$lfstat <- as.character(as_factor(df$lfstat)) 
  return(df)
}

df <- call.cps.org.data()

```  

</div> 

<a id="displayText" href="javascript:toggle(2);">`Stata`</a>  
<div id="toggleText2" style="display: none">  

```{r loading data Stata SHC, eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}
* How to get the data:
use "/Users/fhoces/Documents/data/CPS/cepr_org_2013.dta", clear

*Following the notes here (https://cps.ipums.org/cps/outgoing_rotation_notes.shtml) I generate the weights as orgwgt/12
cap drop *_weight
gen final_weight = orgwgt/12
gen round_weight = round(orgwgt/12, 1)

* There are 1293 cases with missin values for the weigths. I delete them from the data. 
drop if orgwgt == .
sum(orgwgt)
```  

</div> 

### Wage variable
We assume no further adjustments like imputation for top coding, trimming, excluding overtime/commissions, or imputation of usual hours for ''hours vary'' respondents. The CEPR ORG data includes several wage variables ([described here](http://ceprdata.org/cps-uniform-data-extracts/cps-outgoing-rotation-group/)). The wage variable that best matches the description above is `wage3`. This variable measures earnings for hourly workers (excluding overtime, tips, commissions and bonuses -otc-) and non hourly workers (including otc). According to CEPR "...attempts to match the NBER’s recommendation for the most consistent hourly wage series from 1979 to the present"

### Note on CBO's Wage Adjustment
In the original report from CBO, an adjustment was made to the wage of all the workers that did not report an hourly wage (`wage3` is estimated as usual salary per self-reported pay-period over usual hours per pay-period). The goal was to reduce the measurement error in those wages, following the methodology proposed in [this paper](https://www.aeaweb.org/articles?id=10.1257/aer.96.3.461) and compute the adjusted wage as a weighted average of the original wage and the average wage of workers with similar characteristics. 


$$
\begin{equation}
w_{ig} = \alpha w^{raw}_{ig} - (1 - \alpha)  \overline{w^{raw}_{g}} \\
\text{with:    } \quad \overline{w^{raw}_{g}} = \frac{\sum_{g} w^{raw}_{ig} }{N_{g}}  
\label{eq:2}
\tag{2}
\end{equation}
$$ 
  
To implement such adjustments into this dynamic document, we requiere information from CBO on: $\alpha$ and $G$ in $\eqref{eq:2}$.    

### Wage forecast  
To simulate the policy effects we need the distribution of wages and employment under the status quo. From the perspective of 2013, this implies forecasting to 2016 data on employment and wages available in 2013. 

We forecast the wage distribution, from 2013 to 2016 in the following way:

#### Growth adjustments
We assume that the growth forecasts were taken from the 10-Year Economic Projections from CBO ([this website](https://www.cbo.gov/about/products/budget_economic_data)). Annualized growth rates for the number of workers $g_{workers}$, and nominal wage per $g_{wages}$ worker where computed as follows:  

$$
\begin{aligned}
\widehat{ g_{workers} } &= \left[ \frac{\widehat{ N_{workers}^{2016} } }{N_{workers}^{2013}} \right]^{1/3}- 1 \\
\widehat{ g_{wages} }  &= \left[ \frac{\widehat{ Wages^{2016} } / \widehat{ N_{workers}^{2016} } }{Wages^{2013} / N_{workers}^{2013}} \right]^{1/3} - 1 
\end{aligned}
$$ 

The original report assumes higher wage growth for high wages than low wages. To create different rates of growth in wage, we compute different wage growth rates for each decile of wage. The increments across deciles were constant and the set to match a final lowest decile with a yearly growth rate of 2.9%.  

The adjustment over number of workers was made through the weight variable `final_weights` (multiplying it by the growth rate) whereas the `wage3` variable was multiplied by the forecast growth rate of per worker wages.   

##### Code to get economic growth forecasts

<a id="displayText" href="javascript:toggle(3);">`R`</a>  
<div id="toggleText3" style="display: none">      

```{r Getting forecast data, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results = "hide"} 

dwld_and_write_d1 <- function()  {
    # Check if csv file is in local machine. Download if not.
    if ( !("CBO_forec_in2013.csv" %in% dir(path = "./rawdata/") ) ) {
    # download data and read from xls (originally downloaded from CBO, then stored in dataverse for stability)
    download.file(url = "https://www.cbo.gov/sites/default/files/51135-2013-02-EconomicProjections.xls", 
                      destfile = "forec_in2013.xls", mode="wb")
  
    # might require this line in a fresh run: 
    # import from xls to data frame
    trends.df <- rio::import( "forec_in2013.xls" , sheet= "2. Calendar Year")
    
    # Save as csv in raw data. 
    write_csv(trends.df, path = "rawdata/CBO_forec_in2013.csv")
  }
 return(read_csv(file = "rawdata/CBO_forec_in2013.csv"))
}

select_vals <- function(trends.df) {
  # Get column of all projections for 2013: get data from 2012 up to 2019)
  sel.col <- which(trends.df==2012, arr.ind = TRUE)[2] 
  sel.col <- sel.col:(sel.col+7)
  # Get row with all projections for wages and salaries in billions of (nominal) dollars
  # Note: the excel file always contains two rows with the words "wage[s]" and "salar[ies|y]", 
  # we are looking for the second one (corresponding to Wages and Salaries under Income)
  sel.row1 <- unique(
                      apply(trends.df,
                            2, function(x)  grep("Wage.*Salar.*", x ) )
                      )[[2]]
  sel.row1 <- sel.row1[2]
  
  # Get row with all projections for number people employed (in millions)
  sel.row2 <- which(trends.df=="Employment, Civilian, 16 Years or Older (Household Survey)", arr.ind = TRUE)[1]
  # Get Price Index. CBO uses Personal Consumption Expenditures (PCE)
  sel.row3 <- unique(
                      apply(trends.df,
                            2, function(x)  grep("Price Index, Personal Consumption", x ) )
                    )[[2]]

  #Keep only rows and colums identified above
  clean.df <- trends.df[c(sel.row1, sel.row2, sel.row3) , sel.col]
  return(clean.df)
 }  
  
compute_rates <- function(var.df) {
  # Compute growth rates. 
    
  #Labeling and formating
  colnames(var.df) <- 2012:(2012+7)
  var.df <- apply(var.df, 2, as.numeric)
  row.names(var.df) <- c( "wages(total)", "workers",  "Personal CPI")
  
  #Generate wage and non-wage income per worker
  var.df <- rbind( var.df ,  
                      (var.df["wages(total)", ] * 1e9 ) / ( var.df["workers", ] * 1e6) )
  
  # "Price Index, Personal Consumption" = "Personal CPI"
  row.names(var.df) <- c("wages(total)", "workers",  "Personal CPI",
                         "wages per worker")
  # Transpose the data
  var.df <- t(var.df)
  
  # Define a new data set with the anual growth rate of each variable over time
  return( var.df/lag(var.df,1) - 1 )
  }

# Growth rate over several years. This function will be called several times down below. 
gr.factor <- function(var1, init.year, last.year) {
# Compute the compounded growth factor for a given variable in a time interval
# For example growth factor between years 1,2 and 3 will be:
# (1+growth_rate_yr1) * (1+growth_rate_yr2) * (1+growth_rate_yr3)
    if (init.year == 2012) {init.year <- 2013}
    if (is.null(dim( growth.df[as.character(init.year:last.year),] ) )) { 
      rates <- growth.df[as.character(init.year:last.year),] + 1
    } else {
      rates <- apply( growth.df[as.character(init.year:last.year),] + 1, 2, prod )
    }
    return(rates[var1])
}             

trends.df <- dwld_and_write_d1()    
clean.df <- select_vals(trends.df)
growth.df <- compute_rates(clean.df)

#gr.factor("wages per worker", 2014, 2016)
``` 

</div>

<a id="displayText" href="javascript:toggle(4);">`Stata`</a>  
<div id="toggleText4" style="display: none">  

```{r Getting forecast data - Stata HC, eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}
* Anual growth rates (R code to compute rates in commnets):
* ( gr.factor("wages per worker", 2014, 2016) )^(1/3) - 1
scalar wage_gr = 0.04538147
*( gr.factor("workers", 2014, 2016) )^(1/3) - 1
scalar workers_gr = 0.01550989
```  

</div> 

#### ACA adjustments  

In forecasting the wages, the CBO report describes that some wages will be reduced in response to the individual mandate from the Affordable Care Act. Given that no details of the adjustments are described in the report, we do not incorporate them in this dynamic document. 

#### State level minimum wage adjustments {#state-min-wage}  

CBO had to predict the future changes in the state level minimum wages. We use the actual values implemented by each state. The data comes from the Department of Labor ([here](https://www.dol.gov/whd/state/stateMinWageHis.htm)). 

Whenever the predicted wages were below the 2016 federal minimum wage they were replace by it. 

**Important assumption:** when imputing state level min wages, we assume that no effects on employment where incorporated. 

##### Code to get minimum wage values by state 

<a id="displayText" href="javascript:toggle(5);">`R`</a>  
<div id="toggleText5" style="display: none">    

```{r Getting min wage data , eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results = "hide"}
# Minimum wage by state: download, write and read data 
dwld_and_write_d2 <- function() {
  # Check if data is in local machine and download if not.
  if ( !("state_mw.csv" %in% dir(path = "./rawdata/") ) ) {
    fileURL <- "https://www.dol.gov/whd/state/stateMinWageHis.htm"
    #load webpage
    webpage <- read_html(fileURL)
    #identifying all tables in webpage
    tbls <- html_nodes(webpage, "table")
    #extracting tables
    aux1 <- html_table(tbls, fill = TRUE)
    #table 4 and 5 contain the relevant years
    df1 <- aux1[[4]]
    df2 <- aux1[[5]][,1:5]
    df3 <- left_join(df1, df2, "State or other\r\n      jurisdiction")
    colnames(df3) <- c("State", paste("y", 2007:2017, sep = ""))
    write_csv(df3, path = "rawdata/state_mw.csv")
  }
  return(read_csv(file = "rawdata/state_mw.csv"))
}   

# Clean data
clean_d2 <- function(df1 = state_mw_raw) {
  # Check if raw data is newer than clean data
  a <- file.info("./rawdata/state_mw.csv")$ctime
  b <- file.info("./data/state_mw_clean.csv")$ctime
  if (a > b | is.na(b))  {
    # remove states long name
    df1 <- df1 %>% filter(State != "Guam" & State != "Puerto \r\n      Rico" & State != "U.S. \r\n      Virgin Islands" )
    df1$states <- c("Federal","AL","AK","AZ","AR","CA","CO","CT",
                                "DE","FL","GA","HI","ID","IL","IN","IA","KS","KY",
                                "LA","ME","MD","MA","MI","MN","MS","MO",
                                "MT","NE","NV","NH","NJ","NM","NY","NC",
                                "ND","OH","OK","OR","PA","RI","SC","SD",
                                "TN","TX","UT","VT","VA","WA","WV","WI",
                                "WY","DC")
 
    
    df1$states_codes_cps <- c(NA, 63, 94, 86, 71, 93, 84, 16, 51, 59, 58, 95, 82, 
                              33, 32, 42, 47, 61, 72, 11, 52, 14, 34, 41, 64, 43, 
                              81, 46, 88, 12, 22, 85, 21, 56, 44, 31, 73, 92, 23, 
                              15, 57, 45, 62, 74, 87, 13, 54, 91, 55, 35, 83, 53)                                         

    df1 <- df1[,-which(colnames(df1)=="State")]
    df2 <- df1 %>% select(-c("states", "states_codes_cps"))
    
    # removing text from numeric values of MW
    for (year in paste("y", 2007:2017, sep = "")) {
      aux2 <- sapply(df2[, year], function(x) gsub("[^\\d |^\\. |-]+", "", x , perl = TRUE))
      df2[, year] <- unlist(lapply(str_split(aux2, "-"), function(x) max(as.numeric(x))))
      # Some states report a min wage below the federal one, so we replace
      fed_min <- as.numeric(df2[1, year])
      df2[, year] <- replace(df2[, year], 
                             df2[, year] < fed_min | is.na(df2[, year]), 
                             fed_min)    
      }
    
    df2 <- cbind("state" = df1$states,"state_co" = df1$states_codes_cps, df2)
    write_csv(df2, path = "data/state_mw_clean.csv")
  }
  return(read_csv(file = "data/state_mw_clean.csv"))
}

# Get raw data on MW
state_mw_raw <- dwld_and_write_d2()
# Get clean data on MW
state_min_w <- clean_d2()
# Export MW data to Stata
write.dta(state_min_w, "state_min_w.dta")
```

</div>

<a id="displayText" href="javascript:toggle(6);">`Stata`</a>  
<div id="toggleText6" style="display: none">  

```{r Getting min wage data - Stata , eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}
preserve 
  use "/Users/fhocesde/Documents/dissertation/Replication/state_min_w.dta", clear
  sort states 
  tempfile min_wage
  save `min_wage'
restore
```  

</div> 

#### Code to forecast wages and workers 

<a id="displayText" href="javascript:toggle(7);">`R`</a>  
<div id="toggleText7" style="display: none">  

```{r Adjusting wages to 2016 , eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results = "hide"}
#GENERAL NOTE FOR THIS SECTION: the analysis perfomed here is the same as the one with CPS ASEC, so it should be better to wrap it in a function that is called twice. This way I can make sure that the sensitivity analysis works everywhere. 

#Wage adjutsment

#CBO mentions that the lowest 10th percent gets a 2.9% growth in anual wage
#I compute the anualized growth rate of wages and creat 10 bins of wage growth
#starting at 2.4%, then adjust by minimum wages of 2016 and get a anualized 
#growth of 2.9% for the lowest decile. 

#THIS TWO LINES OF CODE ARE DIFFERENT BETWEEN ASEC AND ORG
wage.gr.f <- function(SA.wage.gr = param.wage.gr) {
  ( ( gr.factor("wages per worker", 2014, 2016) )^(1/3) - 1 ) * SA.wage.gr
}

workers.gr.f <- function(SA.worker.gr = param.worker.gr) {
  ( ( gr.factor("workers", 2014, 2016) )^(1/3) - 1 ) * SA.worker.gr
}

# SAME
# Compute the gap between average wage growth and the growth of the lowest decile (assume by CBO)
half.gap.f <- function(SA.wage.gr = param.wage.gr, 
                       SA.base.growth = param.base.growth) {
  wage.gr.f(SA.wage.gr) - SA.base.growth 
}

# 10 wage growth bins starting from lowest assumed value to 2 times the gap computed above 
wage.gr.bins.f <- function( SA.base.growth = param.base.growth,
                            SA.wage.gr = param.wage.gr) {
  bins <- seq(SA.base.growth, wage.gr.f(SA.wage.gr) + 
        half.gap.f(SA.wage.gr,SA.base.growth), length.out = 10)
  return(bins)
}

# CAUTION: DO NOT apply 'ntile()' fn from dplry as is will split ties differently than 'cut()' and results will not
# be comparable to STATA. 

# NOT THE SAME (power of 3 intead of 4)

# Here we adjust min wages
# SAME
wages.final.cps.org.f <- function(df.temp = df, 
                                  wage.gr.bins.temp = wage.gr.bins,
                                  SA.states.raise = param.states.raise,
                                  SA.wages = param.wages) {
    # Compute deciles of wages
    aux.var  <- wtd.quantile(x = df.temp$wage3, 
                             probs = 1:9/10, weights = df.temp$final_weights)
    # Get state min wages for 2016 and 2013
    st.minws <- state_min_w %>% 
      select( c( "state_co","y2013" ,"y2016", "state") )%>% 
      rename(state_la = state, state = state_co)
    # prepare state key for merge
    df.temp <- df.temp %>% 
      mutate(state_la = as.character(as_factor(state)), 
             state = as.numeric(state))  
    # merge with state MW
    df.temp <- left_join(df.temp, st.minws, "state")
    # Create categories of decile, apply wage growth, and increase is new wage is below 2016 state wage
    df.temp <- df.temp %>%
        mutate( "w3.deciles" = cut(wage3, c(0, aux.var, Inf), 
                                  right = TRUE, include.lowest = TRUE) ,  
                "w3.adj1" =  wage3 * ( 1 + wage.gr.bins.temp[w3.deciles] )^3, 
                "wages.final" = ifelse(w3.adj1> y2016 * SA.states.raise,
                                  w3.adj1,
                                  y2016 * SA.states.raise) * SA.wages  ) 
    return(df.temp)
}

wage.gr <- wage.gr.f()
workers.gr <- workers.gr.f()  
half.gap <- half.gap.f()
wage.gr.bins <- wage.gr.bins.f()
df <- wages.final.cps.org.f()

```  

</div>

<a id="displayText" href="javascript:toggle(8);">`Stata`</a>  
<div id="toggleText8" style="display: none">  

```{r Adjusting wages to 2016 - Stata, eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}
* Forecast wages to 2016 : apply diff growth rates per decile (deciles of growth gen in R)
cap drop w3_*
xtile w3_deciles = wage3 [w =final_weight], nq(10)
gen w3_adj1 = wage3 * (1 + 0.02400000)^3 if w3_decile == 1

replace w3_adj1 = wage3 * (1 + 0.02875144)^3 if w3_decile == 2
replace w3_adj1 = wage3 * (1 + 0.03350288)^3 if w3_decile == 3
replace w3_adj1 = wage3 * (1 + 0.03825432)^3 if w3_decile == 4
replace w3_adj1 = wage3 * (1 + 0.04300575)^3 if w3_decile == 5
replace w3_adj1 = wage3 * (1 + 0.04775719)^3 if w3_decile == 6
replace w3_adj1 = wage3 * (1 + 0.05250863)^3 if w3_decile == 7
replace w3_adj1 = wage3 * (1 + 0.05726007)^3 if w3_decile == 8
replace w3_adj1 = wage3 * (1 + 0.06201151)^3 if w3_decile == 9
replace w3_adj1 = wage3 * (1 + 0.06676295)^3 if w3_decile == 10

* Merge with State min data and replace wages below state min in 2016 by it.
decode state, g(state_s)
sort state_s
merge state_s using `min_wage'
* Drop Guam, PRVI, Federal
drop if _m == 2
drop _m

gen w3_adj_min = w3_adj1
replace w3_adj_min = minwage_2016 if w3_adj1 < minwage_2016
```  

</div> 


## Get the $N$   

### Identify the relevant universe  

```{r temp nums for text below, echo=FALSE, }
#This code chunk computes some of the populations of interest. Specifically, the ones needed for the next
# parragraph.  The temp prefix is used as the next code chunk computes all the relevant populations

temp.working.age.pop <- round( sum(df$final_weights, na.rm = TRUE)/1e6, 1)

temp.employed.pop <- round( sum( df$final_weights * 
                                   (df$lfstat == "Employed"), 
                                 na.rm = TRUE)/1e6, 1)
temp.unemployed.pop <- round( sum( df$final_weights * 
                                     (df$lfstat == "Unemployed"), 
                                   na.rm = TRUE)/1e6, 1)
temp.nilf.pop <-  round( sum( df$final_weights * 
                                (df$lfstat == "NILF"), 
                              na.rm = TRUE)/1e6, 1)
temp.salary <- round( with(df, sum(final_weights * 
                                     (empl == 1 &  
                                        (selfinc == 0 & selfemp == 0)), 
                                   na.rm = TRUE))/1e6 , 1)
temp.salary.g1 <- round(with(df, sum(final_weights * 
                                       ( (empl == 1 & 
                                            selfinc == 0 & selfemp == 0) &
                                           (paidhre == 1 | hrsvary != 1 |
                                              is.na(hrsvary) ) &
                                           (wage3==0 | is.na(wage3) ) ) , 
                                     na.rm = TRUE))/1e6, 1)
temp.nohour <- round(with(df,  sum(final_weights * 
                                     (empl == 1 & 
                                        (selfinc == 0 & selfemp == 0) & 
                                        (paidhre == 0 |is.na(paidhre))), 
                                   na.rm = TRUE))/1e6, 1)
temp.nohour.hours.vary <- round(with(df, sum( final_weights * 
                                                (empl == 1 & 
                                                  (selfinc == 0 & selfemp == 0) &
                                                   (paidhre == 0 | is.na(paidhre) ) &
                                                   hrsvary == 1), 
                                              na.rm = TRUE))/1e6, 1)
temp.pop.of.interest <- round(with(df, sum(final_weights * 
                                             (empl == 1 & 
                                                (selfinc ==0 & selfemp == 0) &
                                                ( (paidhre == 0 & hrsvary != 1) |
                                                    paidhre ==1 )  & wage3 != 0),
                                           na.rm = TRUE))/1e6, 1)
```

According to the CPS data, the population of working age in 2013 was `r temp.working.age.pop` million*. Of those, `r temp.employed.pop` million were working, `r temp.unemployed.pop`, were unemployed and `r temp.nilf.pop` were not in the labor force (NILF). 

Among those employed, `r temp.salary` million workers receive a salary (not self employed or self incorporated).  A small number of salary workers (`r temp.salary.g1` million) did not reported any wages and were excluded from the sample. Of the employed salary workers `r temp.nohour` million did not report an hourly wage and it was computed from their reported pay-period divided by the reported hours in such pay-period. However, `r temp.nohour.hours.vary` million workers from this group reported having varying hours. Their wages were not calculated and were also excluded from the sample.  As a result the final number of workers where a rise in the minimum wage can have a direct effect is `r temp.pop.of.interest` million (= `r temp.salary` - `r temp.salary.g1` -  `r temp.nohour.hours.vary`), this is our universe of interest. Figure 1 presents visual representation of all these populations. 

We know compute some descriptive statistics of the labor force in 2013 and the distribution of wages of the universe of interest both in 2013 and the predicted values for 2016. 

Define variable that tags population of interest

#### Statistics and code behind figure 1

<a id="displayText" href="javascript:toggle(9);">`R`</a>   
<div id="toggleText9" style="display: none">   

```{r desc stats2,  eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, error=FALSE, collapse=TRUE}  
# Tag population of interest: 
# Employed & 
#   (not selfincorporated | not selfemployed) & 
#   ( (not paid by the hour & (hours do not vary or missing in hours vary)) | paid by the hour ) &
#   wages are not zero and not missing
get.pop.int <- function(df.temp=df) {
  df.temp <- df.temp %>% mutate("pop_of_int" = (empl == 1 &
               (selfinc == 0 & selfemp == 0) & 
               ( (paidhre == 0 & ( hrsvary != 1 | is.na(hrsvary) )  )  | 
                   paidhre == 1 )  & 
               !(wage3 == 0 | is.na(wage3) ) ) )
  return(df.temp)
}

df <- get.pop.int()

# Tables 1 - 4 where constructed to look at the data. Only table 4 is shown in the final output 

# Table 1: Describing population of interest (weighted and unweighted)
# To compute the new total of workers we multiply the original weigths by the growth rate. 
f_table1 <- function(var_name, title) {
  var1 <- dplyr::enquo(var_name)
  table_1 <- df  %>% 
     summarise("(1) Total" =
                 sum(!!var1, na.rm = TRUE), 
               "(2) Employed" = 
                 sum( !!var1 * (empl == 1), na.rm = TRUE), 
               "(3) Salary (among employed)" = 
                 sum(!!var1 * (empl == 1 &      #Salary worker if 
                 (selfinc == 0 & selfemp == 0))        #not self employed or     
                  , na.rm = TRUE),                     #self incorp.
  
               "(4) Not Paid hourly (among salary)" = 
                 sum(!!var1 * (empl == 1 &       # Not paid hourly if salary and 
                 (selfinc == 0 & selfemp == 0) &        # not paid hourly
                 (paidhre == 0 | is.na(paidhre) )), na.rm = TRUE), 
  
               "(5) Hours Vary (among not paid hourly)" =
                 sum(!!var1 * (empl == 1 &       #Hours vary if not paid hourly and 
                 (selfinc == 0 & selfemp == 0) &        #hours vary
                 (paidhre == 0 | is.na(paidhre) ) & hrsvary == 1), na.rm = TRUE),
  
               "(6) No wage (in (3) but not in (5))" = 
                 sum(!!var1 * ( (empl == 1 & selfinc == 0 & selfemp == 0) & 
                     (paidhre == 1 | hrsvary != 1 | is.na(hrsvary) ) & 
                     (wage3==0 | is.na(wage3) ) ) , na.rm = TRUE), 
               
               "Population of Interest = (3) - (5) - (6)" = 
                sum(!!var1 * (empl == 1 & (selfinc ==0 & selfemp == 0) & 
                      ( (paidhre == 0 & hrsvary != 1) | paidhre ==1 )  & 
                      wage3 != 0) , na.rm = TRUE)
               )

  table_1 <- t(table_1)
  table_1 <- format(table_1, big.mark = ",", digits = 0, scientific = FALSE)
  colnames(table_1) <- title
  return((table_1))
}

# Table 2: summary statistics of wage variable (levels)
#Summary stats 
sum.stas1 <- function(x, wt) {
   c( "mean" = weighted.mean(x,w = wt, na.rm = TRUE),
      "sd" = sqrt( wtd.var(x, weights = wt) ) , 
      "median" = wtd.quantile( x, weights = wt, prob = c(.5)) ,
                wtd.quantile( x, weights = wt, prob = c(.1, .9) ) )
} 
f_table2 <- function() {
  table_2 <- df %>% 
  filter(pop_of_int == 1 & !is.na(wage3)) %>% 
    with(sum.stas1(wage3, final_weights))

  table_2 <- cbind(table_2)
  colnames(table_2) <- "Wage"
  return(table_2)
}

# Table 3: DESCRIBE
f_table3 <- function() {
  table_3 <- df %>% 
    filter(pop_of_int == 1 & !is.na(wage3)) %>% 
      summarise("> $7.5" = weighted.mean(wage3<7.5,w = final_weights), 
                "> $9" = weighted.mean(wage3<9,w = final_weights), 
                "> $10.10" = weighted.mean(wage3<10.10,w = final_weights), 
                "> $13" = weighted.mean(wage3<13,w = final_weights), 
                "> $15" = weighted.mean(wage3<15,w = final_weights) 
                )
  
  table_3 <- t(table_3)
  colnames(table_3) <- "Perc"
  return(table_3)
}
 
# Table 4: DESCRIBE
f_table4 <- function() {
  table_4 <- matrix(NA, 7, 2)
  colnames(table_4)  <- c("2013", "2016: status quo")
  rownames(table_4) <- c("Salary workers", 
                         "Median wage", 
                         "% < 7.5","% < 9", 
                         "% < 10.10", "% < 13", 
                         "% < 15" )
  
  table_4[1,1] <- table_1[7,1]
  table_4[1,2] <- format(sum(df$final_weights[df$pop_of_int==1] * 
                              (1 + workers.gr)^3, 
                            na.rm = TRUE), big.mark=",")
  table_4[2,1] <- table_2[3]
  
  table_4[2,2] <- round( with(df[df$pop_of_int == 1 & !is.na(df$wages.final), ], 
                              wtd.quantile( wages.final, weights = final_weights * 
                              (1 + workers.gr)^3, prob = c(.5) ) ), digits = 2 )
  
  table_4[3:7,1]  <- round(as.matrix(table_3), digits = 2)
  
  aux.1 <- df %>% 
    filter(pop_of_int == 1 & !is.na(wages.final)) %>% 
      summarise("> $7.50" = weighted.mean(wages.final<7.5,
                                          w = final_weights * (1 + workers.gr)^3), 
                "> $9" = weighted.mean(wages.final<9,
                                       w = final_weights * (1 + workers.gr)^3), 
                "> $10.10" = weighted.mean(wages.final<10.10,
                                           w = final_weights * (1 + workers.gr)^3), 
                "> $13" = weighted.mean(wages.final<13,
                                        w = final_weights * (1 + workers.gr)^3), 
                "> $15" = weighted.mean(wages.final<15,
                                        w = final_weights * (1 + workers.gr)^3) 
                )
  
  table_4[3:7,2] <- round( as.matrix(aux.1), digits = 2 )
  return(table_4)
}


  
### Build first treemap
#if (!(length(dev.list()) == 0)) { dev.off() }
#x11()
# NEED TO DESCRIBE
data_for_treemap1 <- function() {
  universe.1 <- df %>% 
          mutate("teen" = ifelse(age<20, "teen", "adult"), 
                 "selfemp_inc" = 1 * (selfemp == 1 | selfinc == 1), 
                 "pop_of_int" = 1 * pop_of_int ) %>%  
          group_by(lfstat, selfemp_inc, teen, pop_of_int)  %>% 
          summarise("total" = sum(final_weights, na.rm = TRUE))
  
  universe.1$selfemp_inc[universe.1$lfstat!="Employed"] = NA 
  universe.1[universe.1$lfstat!="Employed" | universe.1$selfemp_inc==1, c("teen", "pop_of_int")] = NA
  universe.1$selfemp_inc[universe.1$selfemp_inc==0]  <- "salary"
  universe.1$selfemp_inc[universe.1$selfemp_inc==1]  <- "self employed or self incorporated"
  #universe.1$pop_of_int <- with(universe.1, ifelse(pop_of_int==1,"included", "excluded"))
  return(universe.1)
}

# NEED TO DESCRIBE
treemap.1 <- function(){
  invisible(
  treemap(universe.1,
    index=c("lfstat", "selfemp_inc", "teen"),
    vSize=c("total"),
    range = c(7, 15),
    type="index",
    algorithm="pivotSize",
    fontsize.labels = c(12:8),
    border.col = c("#FFFFFF", "#000000","#000000"), 
    aspRatio= 1.5, 
    palette = c("#D3D3D3"), 
    title.legend="number of employees",
    fontface.labels  = c(3,2,1), 
    align.labels=list(c("left", "top"), c("right", "top"), c("right", "bottom") ), 
    bg.labels = 1, 
    title = "Figure 1: Distribution of population of working age in 2013"
  ))
}

table_1 <- cbind(f_table1(final_weights, "N"), 
                 f_table1(!is.na(final_weights), "Unweighted"))
table_2 <- f_table2()
table_3 <- f_table3()
table_4 <- f_table4()
```

</div>

<a id="displayText" href="javascript:toggle(10);">`Stata`</a>  
<div id="toggleText10" style="display: none">  

```{r desc stats2 STATA, eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}

*Population of interest

*Employment categories:
global employed "empl == 1" 
global salary	"empl == 1 & selfinc == 0 & selfemp == 0"
global nhourly	"empl == 1 & selfinc == 0 & selfemp == 0 & (paidhre == 0 | paidhre ==.)"
global hrs_vary "empl == 1 & selfinc == 0 & selfemp == 0 & (paidhre == 0 | paidhre ==.) & hrsvary ==1"


*Tag poppulation of interest: Salary workers that either paid hourly or not paid by the hour but hours not vary, and have a non zero non missing wage
cap drop pop_of
gen pop_of_int = (empl == 1 & (selfinc ==0 & selfemp ==0) & (paidhre ==1 | (paidhre == 0 & hrsvary != 1))  & (wage3 != 0 & wage3 != .) )

matrix table_1 = J(7,2,99)

*1 -Total 
sum final_weight 
noi di "Total sample in CPS ORG"
noi di %14.2f r(sum)
mat table_1[1,1] = r(sum)

count if final_weight!=. 
noi di "Total sample in CPS ORG: unweighted"
noi di %14.2f r(N)
mat table_1[1,2] = r(N)


*2 -Employed
sum final_weight if $employed
noi di "Population Employed"
noi di %14.2f r(sum)
mat table_1[2,1] = r(sum)


count if $employed
noi di "Population Employed: unweighted"
noi di %14.2f r(N)
mat table_1[2,2] = r(N)


*3 -Salaried worker
sum final_weight if $salary
noi di "Salaried workers"
noi di %14.2f r(sum)
local c = r(sum)
mat table_1[3,1] = r(sum)


count if $salary
noi di "Salaried workers: unweighted"
noi di %14.2f r(N)
local c_uw = r(N)
mat table_1[3,2] = r(N)


*4 -Not paid by the hour
sum final_weight if $nhourly
noi di "Salaried workes who are not paid by the hour"
noi di %14.2f r(sum)
mat table_1[4,1] = r(sum)

count if $nhourly
noi di "Salaried workes who are not paid by the hour: unweighted"
noi di %14.2f r(N)
mat table_1[4,2] = r(N)

*5 -Among those who are not paid by the hour: hours vary
sum final_weight if $hrs_vary
noi di "Salaried workes who are not paid by the hour and hour vary"
noi di %14.2f r(sum)
local a = r(sum)
mat table_1[5,1] = r(sum)

count if $hrs_vary
noi di "Salaried workes who are not paid by the hour and hour vary: unweighted"
noi di %14.2f r(N)
local a_uw = r(N)
mat table_1[5,2] = r(N)


*Among those in group 3 but not 5, how many has no wage
sum final_weight if (empl == 1 & selfinc == 0 & selfemp == 0) & (paidhre == 1 | hrsvary != 1) & (wage3==0 | wage3==.)
noi di "Among those in group 3 but not 5, how many has no wage"
noi di %14.2f r(sum)
local b = r(sum)  + `a'
mat table_1[6,1] = r(sum)

count if (empl == 1 & selfinc == 0 & selfemp == 0) & (paidhre == 1 | hrsvary != 1) & (wage3==0 | wage3==.)
noi di "Among those in group 3 but not 5, how many has no wage: unweighted"
noi di %14.2f r(N)
local b_uw = r(N)  + `a_uw'
mat table_1[6,2] = r(N)

*Population of interest: Salary workers minus:
* 	- those workes who are not paid by the hour and hours vary
*	- any additional workers that doesn't have a wage. 
noi di "Population of interest:"
noi di %14.2f  `c' - `b'
mat table_1[7,1] = `c' - `b'

noi di "Population of interest: unweighted"
noi di %14.2f  `c_uw' - `b_uw'
mat table_1[7,2] = `c_uw' - `b_uw'

sum final_weight if pop_of_int == 1 
noi di "Pop. of interest: workers excluded self* and not paid by the hour whose hours vary"
noi di %14.2f r(sum)

* Table_1:
noi di "Table 1"
noi mat list table_1

*Note: Stata cannot produce treemaps/mosaic plots, but numbers in table 1
*should be identical.  

```

</div>

```{r print fig1, echo=FALSE} 
universe.1 <- data_for_treemap1()
aux.1 <- treemap.1()
```

For the universe of interest (employed, salaried, with hourly wages or non varying hours N = `r temp.pop.of.interest` millions), we describe the distribution of hourly wages in 2013 and the forecast values for 2016. Figure 2.

#### Statistics and code behind figure 2

<a id="displayText" href="javascript:toggle(11);">`R`</a>  
<div id="toggleText11" style="display: none">  
  
```{r Dist wage, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results="hide", error=FALSE, collapse=TRUE, fig.show='hold'}
# Plot with distribution of wages under the status quo in 2013 and 2016
two_dist <- function() { 
  df  %>% 
    filter(pop_of_int==1 & wage3<=200)  %>% 
    select(wage3, wages.final, final_weights) %>%
    melt(value.variables=c("wage3", "wages.final") , id="final_weights") %>% 
    mutate("final_weights" = ifelse(variable=="wages.final", 
                                    final_weights * (1 + workers.gr)^3, 
                                    final_weights) ) %>% 
    ggplot() + 
      geom_density(aes(x = value, 
                       fill=variable, 
                       weight = final_weights, 
                       alpha = 1/2, 
                       colour=variable), bw=1, kernel = "gau") + 
      geom_vline(xintercept = c(7.25, 10.10, 11.5), col="blue") +
      coord_cartesian(xlim = c(0,20)) +   
      guides(alpha = "none", colour="none") + 
      scale_x_discrete(limits = c(0,7.5, 10.10, 11.5, 20)) +
      labs(y = NULL, 
           x = "Wage" , 
           title = "Figure 2: Distribution of wages in 2013 and 2016(forecast)")+
      theme(axis.ticks = element_blank(), axis.text.y = element_blank()) +
      theme(legend.justification=c(0,1), 
            legend.position=c(0,1), 
            legend.background = element_rect(fill = "transparent", 
                                             colour = "transparent") )+
      scale_fill_discrete(name=NULL,
                           labels=c("2013", "2016 (Forecast)"))
}
p <- two_dist()
```

</div> 

<a id="displayText" href="javascript:toggle(12);">`Stata`</a>  
<div id="toggleText12" style="display: none">  
  
```{r Dist wage - Stata, eval=FALSE,echo=display_code, engine="stata", engine.path=statapath, comment=""}
*Figure 2
cap drop *_weight_2016
gen final_weight_2016 = final_weight * (1 + wage_gr)^3
gen round_weight_2016 = round(final_weight_2016) 


#delimit ;
twoway 	(kdensity wage3 if pop_of_int == 1 [fweight = round_weight], 
			bwidth(.9) range(0 20)) 
		(kdensity w3_adj_min if pop_of_int == 1 [fweight = round_weight_2016], 
			bwidth(.9) range(0 20)), 
		title(Figure 2: Distribution of wages in 2013 and 2016(forecast)) 
		xline(7.25 10.10 11.5)  
		xlabel(0 "0" 7.5 "7.5" 10.1 "10.10" 11.5 "11.5" 20 "20")
		legend(order(1 "2013" 2 "2016 (Forecast)"))
		yscale(off) 
		xtitle(wage per hour);
#delimit cr
```

</div> 

```{r print fig2, echo=FALSE, warning=FALSE, message=FALSE} 
print(p)
```  


Table 1 below presents more detail statistics for the wage distributions for 2013 and for the forecast wages of 2016.  
  
```{r, eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, collapse=TRUE, fig.show='hold'}
knitr::kable(table_4, caption="Comparison of 2013 and 2016 under the status quo", digits = 1)
```  

Among the population of interest, the employment effects of the minimum wage will be computed separatedley for adults ($age\geq 20$) and teenagers ($16 \leq age < 20$). For this purpose we present the wage distribution for both groups. Figure 3. 

#### Statistics and code behind figure 3

<a id="displayText" href="javascript:toggle(13);">`R`</a>  
<div id="toggleText13" style="display: none">  

```{r Fig 3, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results="hide", error=FALSE, collapse=TRUE, fig.show='hold'} 
#####################
#if (!(length(dev.list()) == 0)) { dev.off() }
#x11
data_for_treemap2 <- function() {
  universe.1 <- df %>% 
          mutate("teen" = ifelse(age<20, "teen", "adult"), 
                 "wage_c" = ifelse(wage3<10.10 & !is.na(wage3), "w < 10.10", "w >= 10.10"), 
                 "selfemp_inc" = 1 * (selfemp == 1 | selfinc == 1) ) %>% 
            group_by(lfstat, selfemp, selfemp_inc, teen, wage_c)  %>% 
              summarise("total" = sum(final_weights, na.rm = TRUE)) 
  
  universe.1$selfemp_inc[universe.1$lfstat!="Employed"] = NA
  universe.1$teen[universe.1$lfstat!="Employed"] = NA
  universe.1$wage_c[universe.1$selfemp_inc==1 | universe.1$lfstat!="Employed"] = NA
  
  universe.1$wage_n <- as.numeric(as.factor(universe.1$wage_c))
  
  universe.1$selfemp_inc[universe.1$selfemp_inc==0]  <- "salary"
  universe.1$selfemp_inc[universe.1$selfemp_inc==1]  <- "self employed or self incorporated"
  
  universe.1$color <-c("#d95f0e", "#fff7bc")[as.factor(universe.1$wage_c)] 
  universe.1$color[is.na(universe.1$color)] <- "#D3D3D3"
  return(universe.1)
}

treemap.2 <- function() {  
  invisible(
    treemap(universe.1,
       index=c("lfstat", "selfemp_inc", "teen", "wage_c"),
       vSize=c("total"),
       vColor = c("color"),
       range = c(7, 15),
       type="color",
       aspRatio= 1.5, 
       algorithm="pivotSize",
       border.col = c("#FFFFFF", "#000000","#000000","#000000"), 
       sortID="-wage_n", 
       fontsize.labels = c(12:8),
       #aspRatio= c(4,3), 
       title.legend="number of employees",
       align.labels=list(c("left", "top"), c("right", "top"), 
                         c("right", "bottom"), c("center", "center")), 
       fontface.labels  = c(3,2,1,1), 
       bg.labels = 1, 
       title = "Figure 3: Figure 1 + proportion of salary workers earning more/less than 10.10",
       lowerbound.cex.labels = 1
    )
  )
}


#TO DO: display number of workers below red bars in reactive fashion. 
#Maybe not even a plot: only a slider with min wage and a
#reactive box with the number of workers that would be below it.
```   

</div>

```{r print fig3, echo=FALSE, warning=FALSE, message=FALSE} 
universe.1 <- data_for_treemap2()
aux.2 <- treemap.2()   
```  

### Identify relevant population  

Given our universe, the next step is to identify the population that would be actually affected if a raise in the minimum wage takes place.  The two relevant populations to define now are the number of low wage workers ($\widehat{ N_{lowwage} }$) and the number of workers that would earn less than the new minimum wage ($\widehat{ N_{w\leq MW^{1}} }$). CBO defines a low wage as one below \$11.5 dollars per hour in 2016, and the proposed value used for the minimum wage is \$10.10 dollars per hour. From now on we separate each of these groups among adults and teenagers following CBO's convention. 

Three adjustments are applied to the population of workers with wages below the new minimum wage:    
  1 - Workers whose earnings are mainly from tips are tagged and a different minimum wage is apply to them (\$2.13).  
  2 - A fraction $\alpha_{ 1 }$ of $\widehat{ N_{ w\leq MW^{1} } }$ is deleted to account for non compliers.  
  3 - A fraction $\alpha_{ 2 }$ of $\widehat{ N_{ w\leq MW^{1} } }$ is deleted to account for workers not subject to the Fair Labor Standards Act.  

After this three adjustments, performed over the relevant population forecast for the year 2016, we obtain the final population.  

#### Tipped workers   
Additional information needed from CBO: which occupations they used to identify tipped workers and clarify the conceptual need to adjust for this population (as the lower min wage only applies if they make *more* than 7.50 an hour).    

Apply different minimum wage to workers who receive more than \$30 in tips. This was applied to 11 occupations (such as waiter, bartender, and hairdresser ~10% if low wage workers)
 
Given that we do not know which 11 categories the report makes reference to, and which variable that defines the categories, we will use the variable `peernuot` to identify tipped workers. This variable overestimates the number of tipped workers (13% as opposed to the 10% mentioned in the report) because it also contains the workers paid overtime or commissions. 

Tipped workers with wages below 7.25 are 1% of the total tipped workers. Non-tipped workers with wages below 7.25 are 1.6% of the total. 

#### Non compliance    

We estimate the *proportion of low wage workers (wage less then 11.50) that earn less than the their state's minimum wage in 2013* as a proxy for non compliers under the new minimum wage in 2016. The original report mentions that it comes up to 12% of the low wage population. 

Following the original report (footnote 25) we will consider salary workers as non compliers only if their wage is lower than the minimum wage, by less than 25 cents for non-tipped workers or 13 cents for tipped workers. 

##### Code to compute percentage of non compliers 

<a id="displayText" href="javascript:toggle(15);">`R`</a>  
<div id="toggleText15" style="display: none">  

```{r Estimating non compliance, eval=TRUE, echo=display_code, warning=FALSE, message=FALSE}
# Percentage of total workers in 2013 that earn less that their states' minimum wage. 
# Universe is defined by "low wage" workers, with wages below $11.5 an hour.  
# variable peernuot seems to be the most appropiate variable to indicate wether or nor receives tips
# 1=YES; 2=NO
# DESCRIBE
f_non_comp_stats <- function(){ 
  df %>%
  select(wage3, state, final_weights, peernuot, pop_of_int, y2013, y2016)  %>%
    filter(pop_of_int == 1 & wage3<=11.5)  %>%
      summarise(
        "% of non compliers w/o adj" = 
          wtd.mean(wage3 < y2013, 
                          weights = final_weights), 
        "% of non compliers with adj" = 
          wtd.mean(wage3 + 0.25 * (peernuot == 2) + 
                          0.13 * (peernuot == 1)  < y2013, 
                          weights = final_weights) 
        )
}

non.comp.stats <- f_non_comp_stats()
```  

</div>


#### Not covered that might benefit   

Additional information needed from CBO: how they identified non-FLSA eligibility.   

 - Include not covered by FLSA but expected to be affected: employees of small firms, occupations generally exempt from FLSA, and teenagers in first 90 days of employment.  

```{r Target population fig MAYBE DELETE, eval=FALSE,echo=FALSE, warning=FALSE, message=FALSE,  error=FALSE, collapse=TRUE, fig.show='hold'} 
library(treemap)

if (!(length(dev.list()) == 0)) { dev.off() }
#x11()
asd <- df %>% 
        mutate("teen" = ifelse(age<20, "teen", "adult"), 
               "wage_c" = ifelse(wages.final<11.5, "Low Wage", "High wage"), 
               "min_wage" = ifelse(wages.final<10.10, "Target Pop", "Think label"), 
               "selfemp_inc" = 1 * (selfemp == 1 | selfinc == 1) ) %>% 
          group_by(lfstat,selfemp, selfemp_inc, teen, wage_c, min_wage)  %>% 
            summarise("total" = sum(final_weights, na.rm = TRUE)) 

asd$selfemp_inc[asd$lfstat!="Employed"] = NA
asd$teen[asd$lfstat!="Employed" | asd$selfemp_inc == 1] = NA
asd$wage_c[asd$lfstat!="Employed"] = NA
asd$min_wage[asd$lfstat!="Employed" | asd$wage_c == "High wage" ] = NA

asd$selfemp_inc[asd$selfemp_inc==0]  <- "salary"
asd$selfemp_inc[asd$selfemp_inc==1]  <- "self employed or self incorporated"

asd$wage_n <- as.numeric(asd$wage_c)

asd$color <-heat.colors( nlevels(as.factor(asd$min_wage)) )[as.factor(asd$min_wage)]
asd$color[is.na(asd$color)] <- "#D3D3D3"

treemap(asd,
 index=c("lfstat", "selfemp_inc", "teen", "wage_c", "min_wage"),
 vSize=c("total"),
 vColor = c("color"),
 range = c(7, 15),
 type="color",
  aspRatio= 1.5, 
 algorithm="pivotSize",
  border.col = c("#FFFFFF", "#000000","#000000","#000000"), 
 sortID="-wage_n", 
 fontsize.labels = c(12:8),
 #aspRatio= c(4,3), 
 palette = c(rev(cm.colors( nlevels(asd$wage_c) )),"#D3D3D3"), 
title.legend="number of employees",
align.labels=list(c("left", "top"), c("right", "top"), c("right", "bottom"), c("center", "center")), 
fontface.labels  = c(3,2,1,1), 
bg.labels = 1
)

```   

The estimated percentage of non-compliance is `r paste(format(100*non.comp.stats[2], digits=3),"%", sep="")` of the target population in 2013 (N = `r table_4[1,2]`). 

### Summary  

Define $\hat{g(2016|2013)}$ as the growth factor for the population from the year 2013 to 2016 ($\hat{g(2016|2013)} = (1 + \hat{g})^3$, where $\hat{g}$ is the annual growth rate of the population). Then:  
<!--
$$ 
\begin{aligned}
\widehat{ N^{final}_{teen} } &= \hat{N_{low \, wage}} (2016|2013)  \times P(\hat{w} \leq MW^{1} | \hat{w} \leq  low \, wage) \times
(1 - \hat{\alpha_{1}} - \hat{\alpha_{2}}) \times P(Age \leq 19 | \hat{w} \leq MW^{1}) \\
&= \hat{g(2016|2013)} \times  \hat{N_{employed}} (2013) \times P(\hat{w} \leq low\,wage)  \times P(\hat{w} \leq MW^{1} | \hat{w} \leq  low \, wage) \times \\
&\quad (1 - \hat{\alpha_{1}} - \hat{\alpha_{2}}) \times P(Age \leq 19 | \hat{w} \leq MW^{1}) \\
&= \hat{g(2016|2013)} \times  \hat{N_{employed}} (2013) P(\hat{w} \leq MW^{1} ) \times (1 - \hat{\alpha_{1}} - \hat{\alpha_{2}}) \times P(Age \leq 19 | \hat{w} \leq MW^{1})
\end{aligned}
$$
-->
$$ 
\begin{aligned}
\widehat{ N^{teen}_{final} } &= \hat{N^{teen}_{\hat{w} \leq MW^{1} } } (2016|2013) \times 
(1 - \hat{ \alpha^{teen}_{1} } - \hat{ \alpha^{teen}_{2} })\\
&= \hat{ g(2016|2013) } \times  \hat{ N^{teen}_{employed} } (2013) \times P(\hat{w} \leq MW^{1}|teen)  \times (1 - \hat{ \alpha^{teen}_{1} } - \hat{ \alpha^{teen}_{2} })
\end{aligned}
$$

Analogously for the adult population: 

$$ 
\begin{aligned}
\widehat{ N^{adult}_{final} } &= \hat{ g(2016|2013) } \times  \hat{ N^{adult}_{employed} } (2013) \times P(\hat{w} \leq MW^{1}|adult)  \times (1 - \hat{ \alpha^{adult}_{1} } - \hat{ \alpha^{adult}_{2} })
\end{aligned}
$$


The table below presents the estimate from 2013 for all each component.

<a id="displayText" href="javascript:toggle(16);">`R`</a>  
<div id="toggleText17" style="display: none">  

```{r compute Ns, eval=TRUE, echo=display_code, warning=FALSE, message=FALSE} 
N.final.f <- function(df.temp = df, workers.gr.temp = workers.gr){
  aux.1 <-  bind_rows( 
    df.temp  %>% 
    filter(pop_of_int == 1) %>% 
    mutate("adult" = ifelse(age>=20, "Adult", "Teen") )  %>% 
    group_by(adult)  %>% 
    summarise( "Salary workers ($\\hat{ N_{employed} }$) (millions)" = sum(final_weights)/1e6,
               "Low wage workers ($w \\leq 11.5 p/h$) (millions)" = sum(final_weights * (wages.final <=11.50))/1e6,
               "% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)" = wtd.mean( 1*(wages.final <=10.10), 
                                         na.rm = TRUE, weights = final_weights) * 100),   
    df.temp %>% 
    filter(pop_of_int == 1) %>% 
    summarise( "Salary workers ($\\hat{ N_{employed} }$) (millions)" = sum(final_weights)/1e6,
                "Low wage workers ($w \\leq 11.5 p/h$) (millions)" = sum(final_weights * (wages.final <=11.50))/1e6,
                "% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)" = wtd.mean( 1*(wages.final <=10.10), 
                                        na.rm = TRUE, weights = final_weights) * 100) %>% 
    mutate(adult = "Total" )
  )
  
  #Non compliance (starting from a different denominator: 'pop_of_int == 1 & wage3 < 11.5')
  aux.2 <- bind_rows(
    df.temp %>%
    select(wage3, age, state, final_weights, peernuot, pop_of_int, y2013)  %>%
    filter(pop_of_int == 1 & wage3 <= 11.5) %>% 
    mutate("adult" = ifelse(age>=20, "Adult", "Teen") )  %>% 
    group_by(adult)  %>% 
        summarise(
            "% of non compliers ($\\alpha_{1}$)" = 
            wtd.mean(1 * (wage3 + 0.25 * (peernuot == 2) + 
                            0.13 * (peernuot == 1)  < y2013 ), 
                            weights = final_weights)*100 ),  
    df.temp %>%
    select(wage3, age, state, final_weights, peernuot, pop_of_int, y2013)  %>%
    filter(pop_of_int == 1 & wage3 <= 11.5) %>% 
        summarise(
            "% of non compliers ($\\alpha_{1}$)" = 
            wtd.mean(1 * (wage3 + 0.25 * (peernuot == 2) + 
                            0.13 * (peernuot == 1)  < y2013 ), 
                            weights = final_weights)*100 ) %>% 
            mutate("adult" = "Total")
  )
  stats2 <- rbind(t(aux.1[,-1 ]), t(aux.2[, -1]) ) 
  colnames(stats2) <- t(aux.1[,1])
  stats2 <- rbind(stats2, "$\\hat{ g(2016|2013) }$" = (1 + workers.gr.temp)^3 )
  aux.total <- apply(stats2, 2, function(x) x[5] * x[1] * (x[3]/100) * (1 - x[4]/100)  )
  return(rbind(stats2, "$\\widehat{ N_{final} }$ (millions)" = aux.total))
}

#RENAME stats2 with table.n.final
table.n.final <- N.final.f()

# FH: There a small difference in the overall total and the sum of Teens and Adults:
# table.n.final[6,3] - sum( (table.n.final[6, 1:2])  )
# I have pinned down the source of the problem to differences in the way % is calculated (for groups relative to the total)
``` 

</div>

```{r table summary ,echo=FALSE, warning=FALSE, message=FALSE} 
knitr::kable(table.n.final, caption="Characteristics of target population", digits = 2)
```  



## Get the $\eta \times \Delta w$  

In order to get the elasticity of labor demand that best fits the context analysed here, two steps were required: (i) identify from the literature the best estimate available; (ii) extrapolate that estimate to fit the specific context of this policy analysis. 

### Getting the best estimates from the literature  

It is unclear the precise mechanism used by CBO to choose the estimates for the labor demand elasticity. 

<!--
Later I will assume that it came from a meta-analysis and calibrate the weights of some of the meta-analysis cited in the report to reflect that choice.  
-->

We take their estimate as given: -0.1 for the teenager population, with a "likely range" a range from 0 to -0.2 ("likely range" is used throughout the report to estimate the expert judgment that the elasticity will be in that range 2/3 of the time)[^1]. The reasons provided in the report for choosing this figure can be summarize by the following three points:  
  - More weight was given to studies that exploit across state variation (as opposed to over time country level variation).    
  - The final estimate takes in to account publication bias towards highly negative estimates.  
  - The magnitude of the increase (%39) and the fact that would be indexed to inflation going forward, makes it an unusually large increase in the minimum wage.    
With this elements we can write the chosen elasticity from the literature ($\eta_{lit}$) as the product of the original assessment of the literature ($\eta_{lit}^{0}$), a reduction factor for publication bias ($F_{pub.bias}<1$) and a amplification factor for a larger variation in minimum wage ($F_{large.variation}>1$):

$$ 
\begin{aligned} 
\eta^{teen}_{lit} &= \eta_{lit}^{0} \times F_{pub.bias} \times F_{large.variation} = -0.1
\end{aligned}
$$ 
  
Additionally, CBO provides two additional caveats that could be added to the analysis:  
 - Ripple effects on employment were assume to be null as the result of two opposing effects: (i) "ripple-wages" would increase unemployment but (ii) substitution of marginally more productive workers for layed off workers below the new minimum wage would decrease unemployment. CBO assumes these effects roughly cancel each other.       
 - CBO acknowledges that effects could be larger (in abs value) during recessions but does not predict a recession for 2016. No estimation is provided of how much larger those effects would be in case of a recession.  
 
### Extrapolating research estimate to current context  
Three adjustments are proposed: (i) extrapolate elasticity estimates for teenager to adults; (ii) re-scale elasticity to population affected by the new minimum wage; (iii) adjust elasticities to reflect average wage variation from and increase in the minimum wage.  

#### Extrapolate from teenagers to adult population  

Most of the estimates from the literate are for teenage population. CBO proposes to extrapolate this estimates to the adult population in the following fashion:
$$
\begin{aligned}
\eta^{adults}_{lit} = \eta^{teens}_{lit} \times F_{extrapolation}
\end{aligned}
$$  

Where a value $F_{extrapolation} = 1/3$ is chosen to reflect that the demand for adult labor is suspected to be more inelastic than the demand for teen labor. 

#### Re-escale elasticities to the population affected by the minimum wage  

The literature reports the estimated effect for a given population $\eta_{lit}$. This estimate can be seen as the weighted average between the demand elasticities for the directly affected population ($\eta_{w\leq MW}$), with wages below the new minimum, and the non affected ($\eta_{w > MW}$) populations, with wages above the new minimum: 

$$
\begin{aligned}
\eta^{g}_{lit} =& p^{g}_{w\leq MW} \eta^{g}_{w\leq MW} + (1 - p^{g}_{w\leq MW})\eta^{g}_{w > MW}  \hspace{4em} g = \{teens, \, adults \}
\end{aligned}
$$ 

The underlying assumption is that $\eta_{w > MW} = 0$. With this the first proposed adjustment becomes: 

$$
\begin{aligned}
\eta^{g}_{w\leq MW} =& \frac{\eta^{g}_{lit}}{p^{g}_{w\leq MW}}  \hspace{4em} g = \{teens, \, adults \}
\end{aligned}
$$ 

The fraction of the population with a forecast income below the new minimum wage ($p^{g}_{w\leq MW}$) is `r paste(round(table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",2], 1),"%", sep="")` for teenagers and `r paste(round(table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1], 1),"%", sep="")` for adults.  

#### Adjust elasticities to average wage variation  

Given that the percentual variation from the old wage to the new minimum wage varies for different levels of wages, total effect should be computed as $\sum_{b} \% \Delta w_{b} \eta^{g}_{w\leq MW} \times N_{b}$ for $b = 1 \dots B$ wage brackets.

The report approximates this calculation by computing the employment effect for average wage variation across the total population, in age group $g$, affected by the minimum wage:  $\overline{\%\Delta w^{g}} \times \eta^{g}_{w\leq MW} \times \widehat{ N^{final}_{g} }$.  

Finally CBO argues that as the variation changes the elasticity should be re-scaled to reflect such variation. With the elasticity resulting in:

$$
\begin{aligned}
\widetilde{ \eta^{g}_{w\leq MW} } &=  \frac{\eta^{g}_{lit}}{p^{g}_{w\leq MW}} \times \frac{\%\Delta MW}{\overline{\%\Delta w^{g}}} = \eta^{g}_{lit} \times F^{g}_{adjs}  \hspace{4em} g = \{teens, \, adults \}
\end{aligned}
$$

Looking at historical trends in the CPS, CBO estimates that $F^{g}_{adjs}$ is 4.5 for both populations[^2].  In the following table we summarize all the elements required to compute $\widetilde{ \eta^{g}_{w\leq MW} }$ 

## Other factors  
CBO reasons that a rise in the minimum wage would have effects in aggregate consumption and this in turn would have effects on employment. The overall effect is estimated as an increase in employment between 30,000 and 50,000 jobs ("a few tens of thousands of jobs"). A narrative argument is provided for the mechanisms behind this effect.

The effects on consumption are separated into direct and indirect. 

### Direct effects on consumption 

 - Job loses $\Rightarrow$ reduction in consumption.  
 - Increase wages $\Rightarrow$ increase consumption.  
 - Less profits for business owners and shareholders  $\Rightarrow$ reduction in consumption.  
 - Increase prices  $\Rightarrow$ reduction in consumption.   

Overall the direct effect on consumption is estimated[?] to be positive due to a higher marginal propensity to consume of the low wage individuals relative to high income ones.   

### Indirect effects on consumption  

 - Increase in consumption $\Rightarrow$ Increase investment in the future  $\Rightarrow$ Increase consumption in the future.  
 - Increase prices of low-wage-intensive items $\Rightarrow$ increase demand in other items  $\Rightarrow$ Bottleneck in other items until firms adjust. 
 
Overall the indirect effect on consumption is estimated to be negative. 

### Overall effect on consumption and its effect on employment  
CBO estimates [?] that the net effect on consumption would be positive and that its effect on employment would be between 30,000 and 50,000 additional jobs for 2016. This effects are estimated for the short run only. The methodology is mention to be similar to the one used to asses the American Recovery and Reinvestment Act ([found here]())

### Prevent double counting  
The estimated elasticities in the literature already account for approximately 10% of the effects through consumption, so the final effect of consumption here is multiplied by 0.9 to prevent double counting.   

$$
\begin{aligned}
\widehat{OF} &=  40,000 \times 0.9
\end{aligned}
$$

## Computing effects on employment  

Putting all these elements together we get: 
$$
\begin{aligned}
 \widehat{ \Delta E } &= \sum_{g\in\{A,T\}} \left( \widehat{ N^{final}_{g} } \times \widetilde{ \eta^{g}_{w\leq MW} }\times \overline{\%\Delta w^{g}}  \right) - \widehat{OF}
\end{aligned}
$$  

### Code to compute each component 

<a id="displayText" href="javascript:toggle(19);">`R`</a>  
<div id="toggleText19" style="display: none">  

```{r employment effect, eval=TRUE, echo=display_code, warning=FALSE, message=FALSE}  
#DESCRIBE
eta.lit.f <- function(SA.eta.lit = param.eta.lit) - 0.1 * SA.eta.lit 

#DESCRIBE
factor.extrap.f <- function(SA.factor.extrap = param.factor.extrap) 1/3 * SA.factor.extrap

#DESCRIBE
final.other.comp <- function(df.temp = df) { 
  eta.lit <- eta.lit.f()
  factor.extrap <- factor.extrap.f()
  stats3 <-  df.temp  %>% 
    filter(pop_of_int == 1) %>% 
    mutate("adult" = ifelse(age>=20, "Adult", "Teen") )  %>% 
    group_by(adult)  %>% 
    summarise("$\\overline{\\%\\Delta w}$" = wtd.mean( ifelse(wages.final <=10.10,
                                              (10.10 - wages.final)/wages.final, NA) , 
                                       na.rm = TRUE, weights = final_weights) * 100 ) 
  
  stats3 <- rbind("$\\eta_{lit}$" = c( eta.lit * factor.extrap , eta.lit ), 
                  "$\\eta_{w \\leq MW'}$" = c( eta.lit * factor.extrap , eta.lit ) / 
                    (table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100),
                  "$F_{adj}$" = c( 4.5, 4.5 ),
                  t(stats3[,-1]) )
  
  aux.1 <- apply(stats3, 2, function(x) x[1] * x[3])
  stats3 <- rbind(stats3, "$\\widetilde{\\eta_{w\\leq MW}}$"= aux.1)
  
  colnames(stats3) <- c("Adult", "Teen")  
  return(stats3)
}


# N final = gr * N_employed * % below min wage * compliers 
# Adj Fact = 1/(% below min wage) * nominal variation/average variation
# Elasticity = elasticities * Extrap * Adj Fact 
# Employmemt effect = N final * Elasticity * Avg variation

# FUNCTION FOR N FINAL
f_n_final <- function (SA.N = param.N, 
                       SA.fract.minwage = param.fract.minwage, 
                       SA.noncomp = param.noncomp) {
  (1 + workers.gr)^3 *
  table.n.final["Salary workers ($\\hat{ N_{employed} }$) (millions)",1:2] *  SA.N *
  table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100 * SA.fract.minwage *
  ( 1 - SA.noncomp * table.n.final["% of non compliers ($\\alpha_{1}$)",1:2]/100 - 0)
}

# FUNCTION FOR ADJ FACTOR
f_adj_fact <- function(SA.fract.minwage = param.fract.minwage, 
                       SA.av.wage.var = param.av.wage.var, 
                       frac.below.mw = table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100, 
                       avg.wg.inc = stats3["$\\overline{\\%\\Delta w}$",1:2]/100 ) {
  new.mw <- 10.10
  nom.wg.inc <- new.mw/7.25 - 1
  res1 <- (1/(frac.below.mw) * SA.fract.minwage) * 
      ( nom.wg.inc / ( avg.wg.inc * SA.av.wage.var ) )
  return(res1)
}

# FUNCTION FOR ELASTICITY FINAL
f_elast <- function(SA.eta.lit = param.eta.lit, 
                    SA.factor.extrap = param.factor.extrap, 
                    SA.F.adj = param.F.adj, 
                    var_adj_fact = adj_fact) {
   c( eta.lit.f(SA.eta.lit) * factor.extrap.f(SA.factor.extrap) , 
     eta.lit.f(SA.eta.lit) ) *
    var_adj_fact * SA.F.adj 
}

f_delta_e <- function(SA.av.wage.var = param.av.wage.var) {
  sum( n_final * elas_final * 
        (stats3["$\\overline{\\%\\Delta w}$",1:2]/100) * 
        SA.av.wage.var )+ 0.05 * 0.9 
}

stats3 <- final.other.comp()
n_final <- f_n_final()
adj_fact <- f_adj_fact(SA.fract.minwage = param.fract.minwage, 
                       SA.av.wage.var = param.av.wage.var,
                       frac.below.mw = table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)","Teen"]/100, 
                       avg.wg.inc = stats3["$\\overline{\\%\\Delta w}$","Teen"]/100 ) #1:2 |"Teen"

adj_fact2 <- f_adj_fact(SA.fract.minwage = param.fract.minwage, 
                       SA.av.wage.var = param.av.wage.var,
                       frac.below.mw = table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100, 
                       avg.wg.inc = stats3["$\\overline{\\%\\Delta w}$",1:2]/100 ) #1:2 |"Teen"

#Report
elas_final <- f_elast(var_adj_fact = 4.5) # adj_fact | 1, 4.5
delta.e1 <- f_delta_e()
# -0.5032739

#Following methodology literally
elas_final <- f_elast(var_adj_fact = adj_fact) # adj_fact | 1, 4.5
delta.e2 <- f_delta_e()

#Following methodology in spirit
elas_final <- f_elast(var_adj_fact = adj_fact2) # adj_fact | 1, 4.5
delta.e3 <- f_delta_e()

#I am not convinced that the adjustment is needed in the first place
elas_final <- f_elast(var_adj_fact = 1) # adj_fact | 1, 4.5
delta.e4 <- f_delta_e()

# "Arbitrary" Sens. Anal to get lower unemp (0.9 * wage in c, 0.37 to 0.5, 1/3 to 1/4)
# delta.e3 <- sum( stats3$N * stats3$`% Empl Below MW`/100 * ( 1 - stats3$`% of non compliers with adj`/100 - 0) *
#        0.9*stats3$`% Mean Wage Inc`/100 * 0.1/(stats3$`% Empl Below MW`/100) *
#        ((0.9*stats3$`% Mean Wage Inc`/100)/0.5) * c(1/4, 1) 
# ) - 0.05 * 0.9

# library(foreign)
# df.ma <- read.dta("C:/Users/fhocesde/Documents/dissertation/meta-analysis/minwage1.dta")
# #x11()
# 
# hist(df.ma$tstatistic, breaks = 300, xlim = c(-8,4))
# 
# abline(v = c(-1.96, -1.6, 1.6, 1.96), col = "red")  

#knitr::kable(stats3, caption="Components of Elasticities", digits = 2) 
```    
</div>


```{r table empl effect, echo=FALSE, warning=FALSE, message=FALSE}
knitr::kable(stats3, caption="Components of Elasticities", digits = 2) 
```    

Using all the components described above we get $\widehat{ \Delta^{-} E } =$ `r round(delta.e1*1000)` thousand jobs. The report however computes $F^{g}_{adjs}$ in a different fashion and gets a value of 4.5 (when computing the values of $F^{g}_{adjs}$  from the table below - as oppose to using historical values - we get $\widehat{ \Delta^{-} E } =$ `r round(delta.e2*1000)` thousand jobs). 

# Distributional effects   

In the first step towards obtaining the policy estimates presented in the [introduction](#introduction) we concluded with a figure of $\widehat{ \Delta^{-} E } =$ `r round(delta.e1*1000)` thousand jobs lost.  We now turn to two additional key quantities: the wage gain among those who get a rise do to the new minimum wage, and the distribution of the losses that pay for that raise. The effect of both quantities is estimated at the level of family income.  

## Computing Family income    

As the unit of interest now is the family and detailed information on income is needed, CBO performs the distributional analysis using a different data set from the Current Population Survey. Instead of the ORG, the following analysis uses the CPS Annual Social and Economic Supplement (ASEC) of March 2013. This data contains income information for the year 2012.  

#### Code to load the data and merge state minimum wages

To the CPS ASEC data we also merge the data on state minimum wages describe in section [2.1.4.3](#state-min-wage)


<a id="displayText" href="javascript:toggle(21);">`R`</a>  
<div id="toggleText21" style="display: none">   

```{r loading data cps_asec, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE}   
# List all non-function objects that I need from here forward

# Call data from CPS ASEC March 2013
call.cps.asec.data <-  function() {
  data_use <- "CPER_ASEC"
  
  # Using CEPR ORG data 
  if (data_use == "CPER_ASEC") {
  # Checking if working directory contains data, download if not. 
    if ( !("cepr_march_2013.dta" %in% dir(path = "./rawdata/")) ) {
    	# create name of file to store data
    	tf <- "cepr_march_2013.zip"
    
    	# download the CPS repwgts zipped file to the local computer
    	download.file(url =  "http://ceprdata.org/wp-content/cps/data/cepr_march_2013.zip", tf , mode = "wb" )
    
    	# unzip the file's contents and store the file name within the temporary directory
    	fn <- unzip(exdir = "./rawdata/", zipfile = tf ,"cepr_march_2013.dta",  overwrite = T )
    }
    df <- haven::read_dta("./rawdata/cepr_march_2013.dta")
  }
  
  df <- tbl_df(df)
  return(df)
}

# Add sensitivity analysis parameters to hours, weeks, weights
add.base.vars <- function(SA.hours = param.hours, 
                          SA.weeks = param.weeks, 
                          SA.N = param.N) {
  df %>% mutate("hrslyr" = hrslyr * SA.hours, 
                "wkslyr" = wkslyr * SA.weeks, 
                "hhwgt" = hhwgt * SA.N)
}

# Merge state min wage info
add_minw <- function(df.temp = df) {
  # Get state min wages for 2016 and 2013
  st.minws <- state_min_w %>% 
    select( c( "state_co","y2013" ,"y2016", "state") )%>% 
    rename(state_la = state, state = state_co)
  # prepare state key for merge
  df <- df %>% 
    mutate(state_la = as.character(as_factor(state)), 
           state = as.numeric(state))  
  # merge with state MW
  df <- left_join(df, st.minws, "state")
  return(df)
}

df <- call.cps.asec.data()
df <- add.base.vars() 
df <- add_minw(df)
```   

</div>

### Computing wages in CPS ASEC 2013  

The hourly wage ($w$) was computed as the ratio of yearly earnings ($y$) and the product of usual number of hours worked in a week ($Hour.per.Week$) and the number of weeks worked in a year ($Weeks.per.Year$). The CPS ASEC data set contains three variables for yearly earnings: `incp_all, incp_ern`, `incp_wag` corresponding to all income, earnings, and wages respectively. We choose `incp_wag`. 

For this data set, the weights used in our analysis will be `hhwgt`.  

$$
\begin{aligned}
\hat{w} = \frac{\hat{y}}{\hat{Hours \, per \, Week} \times \hat{Weeks\,per\,Year}}
\end{aligned}
$$

#### Code for computing wages and descriptive statistics  

<a id="displayText" href="javascript:toggle(23);">`R`</a>  
<div id="toggleText23" style="display: none">  

```{r Descript stats1 cps asec, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE} 
# Tag population of interest FH: (1) Need to deal with 598 NA's and (2) clarify that restrictions here are not 
# exactly the same as CPS ORG.  
f_pop_of_int <- function() { 
  df <- df %>% mutate("pop_of_int" =  1 * (empl == 1 &
                (selfinc == 0 & selfemp == 0) & 
                !(incp_wag == 0 | is.na(incp_wag) ) )
                )
  return(df)
}

# For CPS ASEC I am using the weights hhwgt
# DESCRIBE
f_table_5 <- function() {
  table_5  <- df  %>% 
     summarise("(1) Total" =
                 sum(hhwgt, na.rm = TRUE), 
               "(2) Employed" = 
                 sum( hhwgt * (empl == 1), na.rm = TRUE), 
               "(3) Salary (among employed)" = 
                 sum(hhwgt * (empl == 1 &             #Salary worker if 
                 (selfinc == 0 & selfemp == 0))       #not self employed or     
                  , na.rm = TRUE)                     #self incorp.
              )
  
  table_5 <- t(table_5)
  colnames(table_5) <- "N"
  table_5 <- format(table_5, big.mark = ",", digits = 0, scientific = FALSE)
  return(table_5)
}

#Summary stats of wage
sum.stas1 <- function(x, wt) {
   c( "mean" = weighted.mean(x,w = wt, na.rm = TRUE),
      "sd" = sqrt( wtd.var(x, weights = wt) ) , 
      "median" = wtd.quantile( x, weights = wt, prob = c(.5)) ,
                 wtd.quantile( x, weights = wt, prob = c(.1, .9) ) )
} 

#DESCRIBE
f_table_6 <- function() {
  table_6 <- df %>% 
    filter(pop_of_int == 1 & !is.na(hhwgt)) %>%
      select(incp_wag, hrslyr, wkslyr, hhwgt) 
  
  table_6 <- sapply(table_6[,c("incp_wag", "hrslyr", "wkslyr")],
                    function(x) sum.stas1(x, table_6$hhwgt) )
  table_6 <- cbind(table_6)
  colnames(table_6) <- c("Earnings", "Hours", "weeks") 
  return(table_6)
}

# Compute hourly wages, replace negative vales withs 0's
add.wage.var <- function(df) {
  df$wage <- with(df, incp_wag/(hrslyr * wkslyr) )
  df$wage[df$wage<0]  <- NA
  
  df <- df %>% 
      mutate("hhwgt.2013" = gr.factor("workers", 2012, 2013) *  hhwgt , 
           "wage.2013" = gr.factor("wages per worker", 2012, 2013)  *  wage)
  return(df)
}

#DESCRIBE
f_table_7 <- function() {  
  table_7 <- df %>% 
    filter(pop_of_int == 1 & !is.na(wage)) %>%
      summarise("N" = sum(hhwgt.2013),
                "> $7.5" = weighted.mean(wage.2013<7.5,w = hhwgt.2013), 
                "> $9" = weighted.mean(wage.2013<9,w = hhwgt.2013), 
                "> $10.10" = weighted.mean(wage.2013<10.10,w = hhwgt.2013), 
                "> $13" = weighted.mean(wage.2013<13,w = hhwgt.2013), 
                "> $15" = weighted.mean(wage.2013<15,w = hhwgt.2013) 
                ) 
  
  table_7 <- t(table_7) 
  colnames(table_7) <- "Perc (2013)"  
  return(table_7)
}

df <- f_pop_of_int()
table_5 <- f_table_5()
table_6 <- f_table_6()
df <- add.wage.var(df)
table_7 <- f_table_7()
``` 

</div>

#### Adjusting wages 1 REF EQ ABOVE
As with the CPS ORG, an adjustment for wages is applied. Unlike the previous modification, where the adjustments were over a fraction of the population (those who did not report an hourly wage), our understanding is that CBO adjust the wages of all the population in this case.   

The adjustments follows the following formula: 
$$
\begin{aligned}
w_{ig} &= \alpha w^{raw}_{ig} - (1 - \alpha)  \overline{w^{raw}_{g}} \label{samp.eq1_1} \\
\text{with:    } \quad \overline{w^{raw}_{g}} &= \frac{\sum_{g} w^{raw}_{ig} }{N_{g}} \nonumber 
\end{aligned}
$$ 
  
Additional information needed from CBO: $\alpha$ and $G$ in this case.   

#### Adjusting wages 2  NOT SURE

CBO mentions that "it found far fewer workers who would be directly affected by the change in the minimum wage than it had in its analysis of employment", using the CPS ASEC we get `r paste(round(table_7[ "> $10.10",]*100, 1), "%", sep="")` workers below the 10.10 threshold in 2013, while using the CPS ORG we get`r paste(round(table_3[ "> $10.10",]*100, 1), "%", sep="")` in 2013. 

We assume that this second adjustment are a linear transformation ("mostly by adjusting some workers’ wages up to the minimum wage projected to apply to them in 2016 under current law"):
$$ 
\begin{aligned}
\widetilde{w_{ig}} &= (1  + I(U \geq 0) \times F_{1}) w_{ig}I(g \in G_{1}) + \\
&\quad w_{ig}( 1- I(g \in G_{1}))  \quad \text{with:  } U\sim Uniform(a,b) 
\end{aligned}
$$  


#### Forecasting wage
The wage forecast is the same methodology as in section 2.  This methodology is applied to a different data set (CPS ASEC) and for one additional year (forecasting from 2012 to 2016) than with the CPS ORG data.

##### Code to forecast wages, workers

<a id="displayText" href="javascript:toggle(25);">`R`</a>  
<div id="toggleText25" style="display: none"> 
```{r forecasting cps asec wage, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results = "hide"} 
# MODIFY TO CALL PREVIOUS FUNCTIO
#Wage adjutsment
#CBO mentions that the lowest 10th percent gets a 2.9% growth in anual wage
#I compute the anualized growth rate of wages and creat 10 bins of wage growth
#starting at 2.4%, then adjust by minimum wages of 2016 and get a anualized 
#growth of 2.9% for the lowest decile. 
#THIS TWO LINES OF CODE ARE DIFFERENT BETWEEN ASEC AND ORG
wage.gr.asec.f <- function(SA.wage.gr = param.wage.gr) {
  ( ( gr.factor("wages per worker", 2013, 2016) )^(1/4) - 1 ) * SA.wage.gr
}

workers.gr.asec.f <- function(SA.worker.gr = param.worker.gr) {
  ( ( gr.factor("workers", 2013, 2016) )^(1/4) - 1 ) * SA.worker.gr
}

#SAME
half.gap.asec.f <- function(SA.wage.gr = param.wage.gr, 
                     SA.base.growth = param.base.growth) {
  wage.gr.asec.f(SA.wage.gr) - SA.base.growth 
}

wage.gr.bins.asec.f <- function(SA.base.growth = param.base.growth,
                         SA.wage.gr = param.wage.gr) {
  seq(SA.base.growth, wage.gr.asec.f(SA.wage.gr) + 
        half.gap.asec.f(SA.wage.gr,SA.base.growth), length.out = 10)
}

wage.gr <- wage.gr.asec.f()
workers.gr <- workers.gr.asec.f()  
half.gap <- half.gap.asec.f()
wage.gr.bins <- wage.gr.bins.asec.f()


#WRAP IN FUNCTION AND DESCRIBE
# CAUTION: DO NOT apply 'ntile()' fn from dplry as is will split ties differently than 'cut()' and results will not
# be comparable to STATA. 
#NOT THE SAME (power of 3 intead of 4)

add.wages.1 <- function(df.temp = df, aux.var.temp = aux.var) {
  aux.var  <- wtd.quantile(x = df$wage, probs = 1:9/10,weights = df$hhwgt)
  df.temp %>%
        mutate( wage.deciles = cut(wage, c(0, aux.var.temp, Inf) , 
                                right = TRUE, include.lowest = TRUE) ,  
                wage.adj1 =  wage * ( 1 + wage.gr.bins[wage.deciles] )^4) 
}



# Here we adjust min wages
# SAME
wages.final.asec.org.f <- function(SA.states.raise = param.states.raise,
                                SA.wages = param.wages) {
  with( df, ifelse(wage.adj1> y2016 * SA.states.raise,
                   wage.adj1,
                   y2016 * SA.states.raise) ) * SA.wages
}

df <- add.wages.1()
df$wages.final <- wages.final.asec.org.f()
```    
</div>


##### Statistics and code behind figure 4  

<a id="displayText" href="javascript:toggle(27);">`R`</a>  
<div id="toggleText27" style="display: none">  

```{r desc stats2 in forcast asec wage,  eval=TRUE, echo=display_code, warning=FALSE, message=FALSE}
f_table_8 <- function() {
  table_8 <- matrix(NA, 7, 2)
  colnames(table_8)  <- c("2013", "2016: status quo")
  rownames(table_8) <- c("Salary workers", 
                         "Median wage", 
                         "% < 7.5","% < 9", 
                         "% < 10.10", "% < 13", 
                         "% < 15" )
  # Total in 2013
  table_8[1,1] <- table_5[3]
  # projected total in 2016
  # To compute the new total of workers we multiply the original weigths by the growth rate. 
  table_8[1,2] <-  format(sum(df$hhwgt[df$pop_of_int==1] * 
                              (1 + workers.gr)^4, 
                            na.rm = TRUE), big.mark=",")
  
  # Median wage before projections
  table_8[2,1] <- df %>% 
    filter( (pop_of_int == 1) & !is.na(wage) ) %>% 
    with(wtd.quantile( wage, weights = hhwgt, prob = c(.5) ) ) %>% 
    round(digits = 2)
  
  # Median wage after projections
  table_8[2,1] <- df %>% 
    filter( (pop_of_int == 1) & !is.na(wage) ) %>% 
    with(wtd.quantile( wages.final, weights =  hhwgt * 
                         (1 + workers.gr)^4, prob = c(.5) ) ) %>% 
    round(digits = 2)
   
  # Wage distribution in 2013
  table_8[3:7,1]  <- round(as.matrix(table_7[-1]), digits = 2)

aux.1 <- df %>% 
  filter(pop_of_int == 1 & !is.na(wages.final)) %>% 
    summarise("> $7.50" = weighted.mean(wages.final<7.5,w = hhwgt * (1 + workers.gr)^4), 
              "> $9" = weighted.mean(wages.final<9,w = hhwgt * (1 + workers.gr)^4), 
              "> $10.10" = weighted.mean(wages.final<10.10,w = hhwgt * (1 + workers.gr)^4), 
              "> $13" = weighted.mean(wages.final<13,w = hhwgt * (1 + workers.gr)^4), 
              "> $15" = weighted.mean(wages.final<15,w = hhwgt * (1 + workers.gr)^4) 
              )

  table_8[3:7,2] <- round( as.matrix(aux.1), digits = 2 )
  return(table_8)
}

# Histogram of wages below $20 for 2013 and 2016
p2 <-    df  %>% 
  filter(pop_of_int==1 & wage<=200)  %>% 
  select(wage.2013, wages.final, hhwgt, hhwgt.2013) %>%
  gather(key = variable, value = value, -c(hhwgt.2013, hhwgt) ) %>% 
  mutate("hhwgt" = ifelse(variable=="wages.final", 
                                  hhwgt * (1 + workers.gr)^4 , 
                                  hhwgt.2013 ) ) %>% 
  ggplot() + 
    geom_density(aes(x = value, 
                     fill=variable, 
                     weight = hhwgt, 
                     alpha = 1/2, 
                     colour=variable), bw=1, kernel = "gau") + 
    geom_vline(xintercept = c(7.25, 10.10, 11.5), col="blue") +
    coord_cartesian(xlim = c(0,20)) + 
    scale_x_discrete(limits = c(0,7.5, 10.10, 11.5, 20))  +   
    guides(alpha = "none", colour="none") + 
    labs(y = NULL, 
         x = "Wage" , 
         title = "Figure 4: Distribution of wages in 2013 and 2016(forecast)")+
    theme(axis.ticks = element_blank(), axis.text.y = element_blank()) +
    theme(legend.justification=c(0,1), 
          legend.position=c(0,1), 
          legend.background = element_rect(fill = "transparent", 
                                           colour = "transparent") )+
    scale_fill_discrete(name=NULL,
                         labels=c("2013 (Forecast)", "2016 (Forecast)"))

table_8 <- f_table_8()

# print(p2) 
# knitr::kable(table_8, caption="Comparison of 2013 and 2016 under the status quo", digits = 1)
```  

</div>


```{r plot and table for dist asec, echo=FALSE, warning=FALSE, message=FALSE}
print(p2) 
knitr::kable(table_8, caption="Comparison of 2013 and 2016 under the status quo", digits = 1)
``` 

## Imputing policy effects

### Imputing wage gains  
If the wage is below the proposed new minimum (10.10), we increase that wage up to 10.10 for all the eligible population.  

#### Ripple effects   
CBO applies an additional wage increase for wages that are in a neighborhood up to 50% of the max increase ($+-0.5(10.10 - 7.25) = +- \$1.4$). Thus, the final imputed wage is:

$$
\tilde{w} =
\begin{cases}
MW' + 0.5(w - 7.25) \quad if \quad  w \in [8.7, 10.10) \\
w + 0.5(11.5 - w) \quad if \quad  w \in [10.10, 11.5) \\
MW' \quad o/w
\end{cases}
$$

##### Code for ripple effects   

<a id="displayText" href="javascript:toggle(29);">`R`</a>  
<div id="toggleText29" style="display: none">  

```{r positive effects,  eval=TRUE,echo=display_code, warning=FALSE, message=FALSE}
# Increase population size (apply workers growth rate to all pop)

# Create new wage (after inc in min wage)
# Apply ripple effects
wage.ripple.f <- function(SA.ripple = param.ripple) {
  df %>% 
  mutate( "hhwgt.2016" = hhwgt * (1 + workers.gr)^4 , 
          "below_min" = ifelse(wages.final <= 10.10 & pop_of_int == 1,
                            1, 
                            0),
          "below_min" = ifelse(is.na(below_min),
                            0, 
                            below_min), 
          "new.wage"  = ifelse(wages.final<10.10 & pop_of_int %in% 1, 
                            10.10, 
                            wages.final), 
          "new.wage"  = ifelse(wages.final>10.10 & 
                                 wages.final<SA.ripple["scope_above"] & 
                                 pop_of_int==1,
                               wages.final + SA.ripple["intensity"] * 
                                 (SA.ripple["scope_above"] - wages.final),
                               ifelse(wages.final>SA.ripple["scope_below"] & 
                                        wages.final<=10.10 & 
                                        pop_of_int==1,
                                      10.10 + SA.ripple["intensity"] * 
                                        (wages.final - 7.25),
                                      new.wage
                          ) )
  )
}

df <- wage.ripple.f()

# Get the number of workers whose wage would bebow 10.10 in the status quo (in millions)
N_benes <- sum(df$hhwgt.2016[df$wages.final <= 10.10 & df$pop_of_int==1], na.rm = TRUE)/1e6

# Compute total wage increase (yearly, in billions) -without ripple effects and before destroying jobs-
wage.inc <- with(df[df$below_min == 1 & df$pop_of_int==1, ], 
      sum((10.10 - wages.final) * hhwgt.2016 * hrslyr * wkslyr , na.rm = TRUE) ) / 1e9 

# Total gain with ripple effects but without destroying any jobs  
wage.inc.with.ripple <- df %>% 
      with( sum((new.wage - wages.final) *  
                  hhwgt.2016 * hrslyr * 
                  wkslyr , na.rm = TRUE) ) / 1e9 
```  
</div>


#### Substracting non compliers  
In section 2.2 we estimate that `r paste(round(table.n.final["% of non compliers ($\\alpha_{1}$)", "Total"], 1),"%", sep="")` of workers eligible for a rise would not receive such benefit. To account for this fraction of non compliers replace the same fraction of new wages with what would have receive under the status quo.  


##### Code to substract non compliers  

<a id="displayText" href="javascript:toggle(31);">`R`</a>  
<div id="toggleText31" style="display: none">  

```{r non compliesrs in asec,  eval=TRUE,echo=display_code, warning=FALSE, message=FALSE}
# Apply ripple effects
alpha.1 <- table.n.final["% of non compliers ($\\alpha_{1}$)", "Total"] * param.noncomp /100

#Assign old wages to a % of the population (non-compliers)
set.seed(123)
add.nocomp <- function(df.temp=df, alpha.1.temp = alpha.1) {
  #randomly assign no compliance to alpha.1.temp% of the population
  df.temp %>% mutate("no.comply" = ifelse(pop_of_int==1 & wages.final<11.5,   
                                    ifelse(runif( length(new.wage) )< alpha.1.temp, 
                                           "Not comply", "comply"),
                                    NA), 
  #Assing old wages for all the non-compliers
                "new.wage.nocomp" = ifelse(no.comply %in% "Not comply" , 
                                    wages.final, 
                                    new.wage) 
                                    ) 
}

df <- add.nocomp()

# Total gain with ripple effects but without destroying any jobs and accounting for non-compliance
wage.inc.with.ripple.non.comp <- df %>% 
    #filter(below_min == 0)  %>% 
      with( sum((new.wage.nocomp - wages.final) *  hhwgt.2016 * hrslyr * wkslyr , na.rm = TRUE) ) / 1e9 

# Number of workers with wages below new min, that are eligible to receive wage inc (before job loses)
N_benes_compliance_below_min <-  sum(df$hhwgt.2016[(df$wages.final != 
                                                      df$new.wage.nocomp) & df$below_min==1], 
                                     na.rm = TRUE)/1e6
#
N_benes_compliance_above_min <-  sum(df$hhwgt.2016[(df$wages.final != 
                                                      df$new.wage.nocomp) & df$below_min==0], 
                                     na.rm = TRUE)/1e6
N_benes_compliance<- sum(df$hhwgt.2016[(df$wages.final != 
                                          df$new.wage.nocomp)], 
                         na.rm = TRUE)/1e6

pt <- function(x) round(x, 1)

``` 

</div>

After accounting for non compliance, the total number of workers that are potentially eligible for a raise is `r pt(N_benes_compliance)` million. Of that number `r pt(N_benes_compliance_below_min)` would have had a wage below the new minimum and `r pt(N_benes_compliance_above_min)` would have had a wage above 10.10, but receives a wage increase through ripple effects. We now impute job losses in order to obtain the final number of workers who benefit and lost from a raise in the minimum wage. 


### Imputing job losses  

The imputation above so far is applied to all workers below the minimum wage (and the ripple effects). Now we need to remove the $\widehat{\Delta E} = `r round(delta.e1, 2)`$ million workers by imputing them a wage of 0. CBO chose to not move the wage all the way to 0 but to cut it in half and apply such imputation to $2\widehat{\Delta E}$. When destroying jobs CBO argued that the effect would be heavier on teenagers and low wage adults. We implement this in the following algorithm:

Replace $\tilde{w} = \tilde{w}/2$ if :  
  - $w \in [7.25, 10.10)$    
  - $\frac{exp(m(x))}{1 + exp(m(x))} > Uniform[\theta]$  with $m(x) = \beta_{1} TEEN - \beta_{2} ADULT\times w$ and $\beta_{1}, \beta_2 >0$  
  - Choose $\theta$ to destroy $2\widehat{\Delta E}$ jobs.   
  
**Information needed from CBO**. So far we are destroying jobs uniformly across workers earning less than 10.10  

#### Code to impute job loses

<a id="displayText" href="javascript:toggle(33);">`R`</a>  
<div id="toggleText33" style="display: none">  
```{r job destruction,  eval=TRUE,echo=display_code, warning=FALSE, message=FALSE}  

job.killer <- function(df.temp=df) {
  #Assign half of old wages to a % of the population (non-compliers)
  set.seed(123)
  job.cut.factor <- 2
  frac_to_destroy <- df %>% 
    filter(pop_of_int == 1 & wages.final <10.10) %>% 
    summarise(job.cut.factor * (- delta.e1) * 1e6 / sum(hhwgt.2016))
    
  #randomly assign no compliance to alpha.1.temp% of the population
  df.temp %>% mutate("wage_cut" = ifelse(pop_of_int==1 & wages.final<10.10,   
                                    ifelse(runif(length(new.wage) )< as.numeric(frac_to_destroy), 
                                           "half wage", "same wage"),
                                    NA), 
  # Compute variables for each scenario
  # cut jobs, no wage raise
                "cut.wage" = ifelse(wage_cut %in% "half wage",
                                 wages.final/job.cut.factor, 
                                 wages.final) ,
  # cut jobs relative to sq, wage raise
                "new.wage.final" = ifelse(wage_cut %in% "half wage",
                                       wages.final/job.cut.factor, 
                                       new.wage.nocomp),  
  # cut jobs relative to raise, wage raise
  "new.wage.cut" = ifelse(wage_cut %in% "half wage",
                                     new.wage.nocomp/job.cut.factor, 
                                     new.wage.nocomp)
  )  
  
}

df <- job.killer()

# Compute total wage increase after ripple effects (yearly in billions)
wage.gain.total <-  df %>% 
      summarise( "Total wage gain" =  sum( (new.wage.final - wages.final) * 
                   hhwgt.2016 * hrslyr * wkslyr , na.rm = TRUE) / 1e9 , 
                "Total wage loss" = sum( (wages.final - cut.wage) * 
                   hhwgt.2016 * hrslyr * wkslyr , na.rm = TRUE) / 1e9, 
                "Total gain before JD" =  sum( (new.wage.final - cut.wage) * 
                   hhwgt.2016 * hrslyr * wkslyr , na.rm = TRUE) / 1e9
                )  
#with(df, hist(new.wage.final - wages.final, xlim = c(-10,10) , ylim=c(0,3e3) , breaks = 50) )
#df %>% mutate("wls" = ifelse(new.wage.final > wages.final, "winner", ifelse(new.wage.final == wages.final, "nothing", "loser"))) %>% with(wtd.table(wls, weights = hhwgt.2016))

```  
</div>

So far we get a total wage gain (accounting for job losses) of `r pt(wage.gain.total["Total wage gain"])`
billion dollars (in 2016) and a total wage loss of `r pt(wage.gain.total["Total wage loss"])` billions due to workers loosing their jobs. The funds that cover the wage gains have to come from either less profits for business or higher prices for consumers. In the next section we review the distribution of wage gains, wage losses and balance losses across the income distribution. 

## Computing family income under status quo and minimum wage increase  

The family income forecast was computed as the sum of forecast wages and non-wage income:
$$
\begin{aligned}
\widehat{Y_{h}(2016|2013)} &= \sum_{i \in h} \left( g_{w}\hat{w} + \sum_{l}(g_{nw_{l}}\hat{nw_{l}}) \right) 
\end{aligned}
$$ 

The other components of family income were forecast as follows: when a growth rate was available for the sub component it was applied (the only one mentioned is interest and dividends), otherwise the growth rate was equal to the change in the price index for personal consumption 
 
Additional information needed from CBO: how do they decompose the income. 

### Growth of non working population  
Forecasts of population growth were the same for working population. For non working population, the growth rate was matched to CBO forecasts for that group. 

### Other income losses.   
Income losses from a reduction in profits ($\Delta^{-}\pi$) and an increase in aggregate prices ($\Delta^{+}P$) is estimated[?] to be distributed as following: 1% of the losses for those below the poverty line (PL), 29% for those between 1 and 6 PL, and 70% for those above 6PL. 

## Other considerations  

### Economywide income effect  

CBO argues that the overall effect on the economy is positive and of \$2 billion dollars for 2016.  

### Quantifying loses  
- Mix gains and loses. 
- Output lost - increase in aggregate demand (!) 
- net gain (!) of 2 billion dollars.  

### Distributional effects   
- Only results, no methodology at all!  This is probably the most important (and overlooked part of the report)

### Interaction with other programs 
No interactions with other programs is assumed (SNAP or EITC).     


### What is in the 2/3  
  - CBO acknowledges uncertainty in the estimates of elasticity due to possible technological changes in the future.  

# Results 


<a id="displayText" href="javascript:toggle(35);">`R`</a>  
<div id="toggleText35" style="display: none">  

```{r income,  eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results='hide'}
#FOR THE FIRST TIME WE NOW LOOK AT OVERALL INCOME AND THE WHOLE POPULATION
#FH:again, I would rather use the non wage growth

non.wage.gr.f <- function(SA.nonwage.gr = param.nonwage.gr) {
  ( ( gr.factor( "Personal CPI", 2013, 2016) )^(1/4) - 1 ) *
               SA.nonwage.gr
}

#Adjust non-wage income  
#Separate HH income in to wage and non-wage
all.income.f <- function(df.temp = df, non.wage.gr.temp = non.wage.gr) {
  df.temp$non.wage <- (df.temp$incp_ern - df.temp$incp_wag) 
  df.temp$non.wage.2016 <- df.temp$non.wage * ( 1 + non.wage.gr.temp)^4
  
  df.temp <- df.temp %>% mutate("year.wage.1" = new.wage.final * (hrslyr * wkslyr) ) %>% 
        group_by(hhseq) %>% mutate("hhld.wage" = sum(year.wage.1, na.rm = TRUE),
                                   "hhld.non.wage" = sum(non.wage.2016, na.rm = TRUE), 
                                   "hhld.income" = hhld.wage + hhld.non.wage, 
                                   "N.fam"= n(), 
                                    "new.inc.pc" = hhld.income/N.fam) %>% 
                                    select(-(hhld.wage:N.fam) )
    
  df.temp <- df.temp %>% mutate("year.wage.2" = wages.final * (hrslyr * wkslyr) ) %>% 
        group_by(hhseq) %>% mutate("hhld.wage" = sum(year.wage.2, na.rm = TRUE),
                                   "hhld.non.wage" = sum(non.wage.2016, na.rm = TRUE), 
                                   "hhld.income" = hhld.wage + hhld.non.wage, 
                                   "N.fam"= n(), 
                                    "sq.inc.pc" = hhld.income/N.fam) %>% 
                                    select(-(hhld.wage:N.fam) )
  
  df.temp$winners = with(df.temp, ifelse(new.inc.pc >= sq.inc.pc, 
                               new.inc.pc - sq.inc.pc, 
                               0) )
  df.temp$losers = with(df.temp, - ifelse(new.inc.pc <= sq.inc.pc, 
                               new.inc.pc - sq.inc.pc, 
                               0) )
  return(df.temp)
}

non.wage.gr <- non.wage.gr.f()
df <- all.income.f()

# Computing differences in income 
# df %>% 
#  with(sum( (new.inc.pc - sq.inc.pc) * hhwgt.2016, na.rm = TRUE) ) /1e9
# Per capita agregate income has slightly less gains, probably due to hhld with winners and losers. 
# wage.gain.total["Total wage gain"]

losses <- with(df, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) )  - param.net.benef
pop.dist <- wtd.table( with(df, findInterval(x = sq.inc.pc, 
                                             vec = c(-Inf,11740, 6*11740, Inf)) ) , 
                       weights = df$hhwgt.2016 )$sum.of.weights
  
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist
  
win.loss.f <- function(SA.factor.1 = param.factor.1, 
                       SA.net.benef = param.net.benef, losses.pc.temp = losses.pc, 
                       df.temp = df) {
  # Asssing per capita balance losses to each indiv.
  df.temp$balance.loss <- as.numeric(losses.pc.temp[with(df.temp, findInterval(x = sq.inc.pc, 
                                                 vec = c(-Inf,11740, 6*11740, Inf) )) ]) 
  # Build quintiles of income based on status quo income
  bins <- with(df.temp,wtd.quantile(x = sq.inc.pc, probs = 1:4/5, weights = hhwgt.2016))
  df.temp$income.group <- with(df.temp, findInterval(x = sq.inc.pc, vec =  c(-Inf,bins) ))

  #df$income.group <- with(df, findInterval(x = sq.inc.pc, vec =  bins ))
  #classify income into 1, 3 and 6 povery lines
  df.temp$income.group.1 <- with(df.temp, findInterval(x = sq.inc.pc, vec =  c(-Inf,11740*c(1,3,6),Inf)  ) )
  return(df.temp)
}

df <- win.loss.f()

# Compute variation by hhld - plot all the effects 
final_fig1 <- df %>% select(new.inc.pc,sq.inc.pc, hhwgt.2016, balance.loss)  %>% 
  mutate( "variation" = new.inc.pc - sq.inc.pc, 
          "sixthtile" = findInterval(x = sq.inc.pc, 
                                     vec = c(-Inf,11740, 11740*1:6, Inf) )) %>% 
  ggplot(aes(sq.inc.pc, variation))  + 
  geom_hline(color = "gray", alpha = 0.5, yintercept = 0, size = 1) +

  geom_jitter( colour = "black", alpha = 1/30, size = 1/10, 
               position = position_jitter(height = 30, width = 15) ) + 
  coord_cartesian(xlim = c(0,2e5), 
                    ylim = c(-5e3,5e3) )+
  theme_minimal() + 
  labs(y = "Change relative to no raise in min wage", 
       x = "Per capita income" , 
       title = "Distribution of gains and losses across per capita income") +
  geom_vline(xintercept = 11740*1:6, 
             col="red", size = 1/3) +
  scale_y_discrete(limits = c(-50e2, -25e2, 0,25e2, 50e2), 
                   labels = c("-5K", "-2.5K", "0", "2.5K", "5K")) + 
  scale_x_discrete(limits = c(5e3,50e3, 103e3, 150e3), 
                   labels = c("5K","50K", "100K", "150K")) +
  geom_abline(color = "red", alpha = 0.2, slope=-1/2, intercept=0,
  na.rm = FALSE, show.legend = NA) + 
  geom_abline(color = "blue", alpha = 0.2, slope=(10.10/7.25 - 1), intercept=0,
  na.rm = FALSE, show.legend = NA) +
  geom_jitter(aes( sq.inc.pc, - balance.loss - 120 ), 
              colour = "blue", alpha = 1/30, 
              size = 1/10, 
              position = position_jitter(height = 30, width = 15) )  +
  geom_text(x = 15e3, y = -4e3,angle = 0, 
            label = c("1PL"), 
            size = 3, colour = "red", data = data.frame()) + 
  geom_text(x = 12e3*6+3e3, y = -4e3,angle = 0, 
            label = c("6PL"), 
            size = 3, colour = "red",  data = data.frame()) +
  theme(plot.title = element_text(hjust = 0.5))
#print(final_fig1)
#ggsave("alt_pe1.png")

# Compute variation by hhld - plot all the effects in same units (average per group)
final_fig2 <- df %>% 
  select(new.inc.pc,sq.inc.pc, hhwgt.2016, balance.loss, winners, losers)  %>%   
  mutate( "variation" = new.inc.pc - sq.inc.pc, 
          "sixthtile" = findInterval(x = sq.inc.pc, 
                                     vec = c(-Inf,11740*1:6, Inf) ) ) %>%   
  filter(sixthtile>=0) %>% 
  group_by(sixthtile) %>% 
  summarise("mean" = wtd.mean(variation, weights = hhwgt.2016),
            "mean (win)" = wtd.mean(winners, weights = hhwgt.2016), 
            "mean (lose)" = wtd.mean(losers, weights = hhwgt.2016),
            "mean (bal lose)" = wtd.mean(balance.loss, weights = hhwgt.2016),
            "N" = sum( hhwgt.2016)   
            ) %>% 
  select(`mean (win)`, `mean (lose)`, `mean (bal lose)`, sixthtile) %>% 
  melt(value.variables=c("mean (win)", "mean (lose)") , id="sixthtile") %>% 
  ggplot(aes(as.factor(sixthtile), value, fill = as.factor(variable))) + 
  geom_bar(color = "gray", alpha = 0.5, 
           stat = "summary", fun.y = "mean", 
           position = "dodge") +
  coord_cartesian(ylim = c(0,300) ) +
  geom_text(x = 6.8, y = 265,angle = 90, label = - round(losses.pc[3]) , size = 3, colour = "#6699FF", alpha = 0.5) + 
  geom_segment(aes(x = 7, y = 200, xend = 7, yend = 310), 
               colour = "#6699FF", arrow = arrow(length = unit(0.2, "cm"))) + 
  theme(legend.justification=c(0, 0), 
                legend.position=c(0, .7),
                legend.background = element_rect(colour = 'transparent', fill = 'transparent') 
                ) +
  scale_fill_discrete(name=NULL, 
                      labels=c("Wage +", "Wage -", "Balance -")) + 
  labs(y = "$/year", 
       x = "Poverty Lines" , 
       title = "Distribution of gains and losses across poverty lines") +
  theme(plot.title = element_text(hjust = 0.5)) 
 #print(final_fig2) 
 #ggsave("final_fig2.png")

quintiles <- with(df,wtd.quantile(x = sq.inc.pc, probs = 1:4/5, weights = hhwgt.2016))
df$income.group <- with(df, findInterval(x = sq.inc.pc, vec =  c(-Inf,quintiles) ))

# Compute variation by hhld - plot all the effects in same units (average per group) with quintiles instead of poverty lines.
final_fig3 <- df %>% 
  select(new.inc.pc,sq.inc.pc, hhwgt.2016, balance.loss, winners, losers, income.group)  %>%   
  mutate( "variation" = new.inc.pc - sq.inc.pc, 
           "inc_gain" = ifelse(variation>0, 
                              "gain", 
                              ifelse(variation < 0, "loss", "same" ) ) ) %>% group_by(income.group) %>% 
  summarise("mean" = wtd.mean(variation, weights = hhwgt.2016),
            "mean (win)" = wtd.mean(winners, weights = hhwgt.2016), 
            "mean (lose)" = wtd.mean(losers, weights = hhwgt.2016),
            "mean (bal lose)" = wtd.mean(balance.loss, weights = hhwgt.2016),
            "N" = sum( hhwgt.2016)   
            ) 

high.loss <-  as.numeric(final_fig3[max(final_fig3$income.group), "mean (bal lose)"])

#THIS IS THE PLOT THAT I WANT TO OUPUT TO THE SHINY APP
final_fig3 <- final_fig3 %>% 
  select(`mean (win)`, `mean (lose)`, `mean (bal lose)`, income.group) %>% 
  melt(value.variables=c("mean (win)", "mean (lose)") , id="income.group") %>% 
  ggplot(aes(as.factor(income.group), value, fill = as.factor(variable))) + 
  geom_bar(color = "gray", alpha = 0.5, 
           stat = "summary", fun.y = "mean", 
           position = "dodge") +
  coord_cartesian(ylim = c(0,400) ) +
  geom_text(x = 4.8, y = 365,angle = 90, 
            label = - round(high.loss) , 
            size = 3, colour = "#6699FF", alpha = 0.5) + 
  geom_segment(aes(x = 5, y = 300, xend = 5, yend = 410), 
               colour = "#6699FF", arrow = arrow(length = unit(0.2, "cm"))) + 
  theme(legend.justification=c(0, 0), 
                legend.position=c(0, .7),
                legend.background = element_rect(colour = 'transparent', fill = 'transparent') 
                ) +
  scale_fill_discrete(name=NULL, 
                      labels=c("Wage +", "Wage -", "Balance -")) + 
  labs(y = "$/year", 
       x = "Quintiles of per capita income" , 
       title = "Distribution of gains and losses across quintiles") +
  theme(plot.title = element_text(hjust = 0.5))   

```

```{r, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results='hide'}
# THIS IS THE PLOT THAT I OUTPUT FOR THE PAPER (AND THE INITIAL SA)
# NEED TO MAKE SURE I AM EXPORTING IN SAME DIMENSIONS
# print(final_fig3)  
# ggsave("policy_est.png")
# 5.89 x 3.69 in image
#NEED TO CHECk WHY RIPPLE EFFECTS AFFECT WAGE LOSES
```


```{r, eval=TRUE,echo=display_code, warning=FALSE, message=FALSE, results='hide'}
# Sample decision: value cost an benefits equally and distribution 1/q 
sum( (df$winners - df$losers - df$balance.loss) * df$hhwgt.2016 )

#need to change 3 for median below
dist.pref.f  <- function(rho, qi, qt) (1 - rho*(qi - 3))/sum(1 - rho*(1:qt - 3)) * qt

 # sum cost a and benefits at the individual level
final_dec.f <- function(x, df.temp = df) {
  sum( (df.temp$winners - df.temp$losers - df.temp$balance.loss) * df.temp$hhwgt.2016 * 
         dist.pref.f( rho = x, qi = df.temp$income.group, qt = max(df.temp$income.group)) ) /1e9
}
# weigth final CBA by quintile



#abline(v = inc.quartiles, col = "blue", lty =1, cex=3)



table_9 <- matrix(NA, ncol = 4, nrow = 5)
colnames(table_9)  <- c("<1PL", "[1PL, 3PL)", "[3PL, 6PL)", ">6PL")
rownames(table_9)  <- c("wage gains", "wage loses", "other loses",
                        "aggr effect", "N_i")

table_9["N_i", ] <- wtd.table(with(df, income.group.1) , weights = df$hhwgt.2016)[[2]]/1e6
  
aux.1 <- df %>% group_by(income.group.1)  %>% summarise(sum((winners) * hhwgt.2016, na.rm = TRUE)/1e9)
table_9[ "wage gains", ] <- t(as.data.frame(aux.1[,2]) )
  
  
#Wage loses: compared to SQ
aux.1 <- df %>% group_by(income.group.1)  %>% summarise(sum((losers) * hhwgt.2016, na.rm = TRUE)/1e9)
table_9[ "wage loses", ] <- t(as.data.frame(aux.1[,2]) )

#Imputing the balnce of the losses
table_9[ "other loses", ] <- as.numeric(losses) * c(0.01, rep(.29/2, 2), 0.70)/1e9

aux.1 <- df %>% 
  group_by(income.group.1)  %>% 
    summarise(sum((winners - losers - balance.loss) * 
                    hhwgt.2016, na.rm = TRUE)/1e9)
table_9[ "aggr effect", ] <- t(as.data.frame(aux.1[,2]) )

#FH: questions:
#    - what is the precise way to define family in CPS?
#    - what is the standard way to choose one obs per family?
```  

</div>




```{r SA1, eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE}
print(final_fig1)
print(final_fig2)
print(final_fig3)
```  

<!-- 

We combine wage gains ($\Delta^{+} W_{i}$), wage loses ($\Delta^{-}W_{i}$) and balance loses ($BL_{i}$) at the individual level to obtain a final effect ($FE_{i}$):

$$
\begin{aligned}
FE_{i} &= \omega_1 \Delta^{+} W_{i} + \omega_2 \Delta^{-} W_{i} + (1 - \omega_1 - \omega_2 ) BL_{i} 
\end{aligned}
$$

Where $\omega_1, \omega_2$ represente the subjective valuation of each effect from the perspective of the policy maker. When aggregating all the final effects the policy maker also will have to make a subjective valuation regarding the redistribution of weatlth. 

$$
\begin{aligned}
FE &= \sum_{i \in N} \frac{ FE_{i} }{D(y_i)^\rho} 
\end{aligned}
$$

Where $D(y_i)$ is the rank function for per capita income $y_{i}$ that returns the decile of income (1 the lowest and 10 the highest), and $\rho$ parametrizes the preferences towards redistribution ($\rho<0$ dislikes redistribution, $\rho>0$ likes redistribution)

```{r SA2, eval=FALSE,echo=FALSE, warning=FALSE, message=FALSE}
quartz()
df <- df.final
png("sample_pref.png", width=2*440, height=2*350, res = 100)
par(oma = c(0,0,0,0),  mar = c( c(2,2,1,2) + .1))
plot(seq(-1/4,1/5,by = 0.01),sapply(seq(-1/4,1/5,by = 0.01) ,final_dec.f), type = "l", main = NULL, axes=F, xlab="", ylab="", xlim = c(-0.2, .2))
axis(side=1, at=c(-0.1), labels=c("Dislikes"))
axis(side=1, at=c(0.1), labels=c("Likes"))
axis(side=1, at=c(0.18), labels=c("Redistribution \n pref.") , pos=-1, tick=F, outer=TRUE, cex.axis=1.1)
  axis(side=1, at=c(0.21), labels=c(expression( (rho) ) ) , pos=-1.5, tick=F, outer=TRUE, cex.axis=1.3)
 
axis(side=1, at=c(-.2, -.1, .1), labels=c(), pos=1, tick=F)

axis(side=2, at=final_dec.f(0.1), labels=c(round(final_dec.f(0.1), 1)) , pos=0.01, tick =F, las = 1)
axis(side=2, at=final_dec.f(-0.1), labels=c(round(final_dec.f(-0.1), 1)) , pos=.04, tick =F, las = 1)
axis(side=2, at=2, labels=c("2") , pos=0, tick =T, las = 1)
axis(side=2, at=0, labels=c("(0,0)") , pos=0.3, tick =F, las = 1)
axis(side=2, at=final_dec.f(0.2), labels=c("W ($bns)") , pos=0.01, tick =F, las = 1, cex = 1.1)


#xlab = "rho", ylab = "W (billions of $)  "

abline(h=0, v=0, lty = 3)
segments(x0 = c(-0.1, 0, 0.1, 0), x1 = c(-0.1, -0.1, 0.1, 0.1 ) , 
         y0 = c(0,final_dec.f(-0.1), 0,final_dec.f(0.1) ), 
         y1 = c(final_dec.f(-0.1), final_dec.f(-0.1), final_dec.f(0.1), final_dec.f(0.1)  ), col = "red", lty = 1)
dev.off()
```  

--> 

<div class="jumbotron">
  <p>Final replication output </p>  
```{r final results, eval=TRUE,echo=FALSE, warning=FALSE, message=FALSE} 

output.template1.f <- cbind(output.template1, "Replication" = c(
  round(sum(table_9["wage gains",]), 1),  
  round(sum(table_9["wage loses",]), 1), 
  round(sum(table_9["other loses",]), 1),  
  2,  
  paste(round(N_benes_compliance, 1),round(N_benes_compliance_below_min, 1), sep = "/"  ), 
  round(-delta.e1,1) )
  )  



table_9[ "wage gains", ] <- t(as.data.frame(aux.1[,2]) )
  
  
#Wage loses: compared to SQ
#aux.1 <- df %>% group_by(income.group)  %>% summarise(mean((sq.inc.pc - cut.income.pc), na.rm = TRUE))  
#table_9[ "wage loses", ] <- t(as.data.frame(aux.1[,2]) )

#Imputing the balnce of the losses
table_9[ "other loses", ] <- as.numeric(losses) * c(0.01, rep(.29/2, 2), 0.70)/1e9


output.template2.f <- cbind(t(output.template2), "Replication loses" =  -round((table_9[ "other loses", ]), 1) 
                          ,"Replication NE" = round(table_9[ "aggr effect", ], 1) )

knitr::kable(
  list(
    output.template1.f,
    t(output.template2.f)
  ),
  caption = 'Policy estimates in CBO report and Replication Results', booktabs = TRUE, 
  align = 'c'
)

#mod.output.template

```   
  <p><a class="btn btn-primary btn-lg" href="https://github.com/fhoces/dd_mw">Learn more</a></p>
</div>


<!-- 
# Extensions 

```{r extensions, }
knitr::kable(round(table_9), caption="Summary of effect of raising the minimum wage", digits = 1)
mod.output.template
```  

--> 

```{r sa, final results, eval=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}


#RUN CODE FROM DD

# 121 * 1.05 * 0.144  * .37 * (1 - .13) * 1/3 * 0.1 + 
#   4 * 1.05 * 0.74   * .37 * (1 - .18) * 0.1
# Remove all data and keep all the functions
rm(list = setdiff(ls(), lsf.str()))

v0 <-  c(-5.328055,  9.328055)
#-5.328055  9.328055
#v0 <-  c( -5.677396,  9.677396)

#step 0: call the data: CPS, growth data, min wage data and CPS ASEC data
#number of salary workers in 2013 (need to improve as is doing diff stuff to ORG and ASEC)
param.N <- 1.0
param.hours <- 1.0
param.weeks <- 1.0
param.states.raise <- 1

# load cps asec data
df <- call.cps.asec.data()
df_asec <- add.base.vars()

rm(df)

df_csporg <- call.cps.org.data()
df <- df_csporg
growth.df <- get.gr.data()
min.wage.data <- readRDS("minwage")
st.minw <- state.minw("2013")
st.minw.2016 <- state.minw("2015")
st.minw.2016[c("AK"	, "AZ", "CA", "CT", "DC"	, "HI"	, "IL", "MA", "MI"	, 
       "MN"	, "MT" , "NV" , "NE" , "NY"	, "OH"	, "RI", "VT"), ] <- 
          c(  9.75	, 8.05, 10	, 9.6, 10.5	  , 8.5	  , 8.25, 10	, 8.5	  , 
      9		  , 8.05 , 8.25 , 9		 , 9		, 8.1	  , 9.6 , 9.6)
colnames(st.minw.2016) <- "2016"

start <- Sys.time()

SA.part1 <- function(
  param.wage.gr = 1.0, #anual average wage growth
  param.worker.gr = 1.0, #anual average workers growth
  param.fract.minwage = 1.0, #Fraction of workers earning below the min wage (in ORG only so far)
  param.av.wage.var = 1.0, #Average wage variation for those affected by min wage (only ORG)
  param.wages = 1.0, # Wages of all workers
  param.nonwage.gr = 1.0, 
  param.eta.lit = 1.0, #Elasticity of labor demand for teenagers
  param.ripple = c("scope_below" = 8.7*1.0, "scope_above" = 11.5*1.0, "intensity" = 0.5*1.0),
  param.base.growth = 0.024 * 1.0, #wage growth of lowest decile (before state min wage increase)
  param.factor.extrap = 1.0, #extrapolation factor from teens to adults
  param.noncomp = 1.0, #Rate of non-compliance
  param.F.adj = 1.0, #Adjustment factor for 'appropiate' population
  param.factor.1 = 1.0,
  param.net.benef = 2e9*1.0,
  param.dist.loss = c(0.01, 0.29, 0.70), #c(0.39567218, 0.53851506, 0.06581275)
  param.jobcut = 1){
    # step 1: compute effect on employment and non-copliance
    #scalar
    wage.gr <- wage.gr.f(SA.wage.gr = param.wage.gr)
    #scalar
    workers.gr <- workers.gr.f(SA.worker.gr = param.worker.gr)  
  
    #scalar
    half.gap <- half.gap.f(SA.wage.gr = param.wage.gr, 
                           SA.base.growth = param.base.growth)
    
    #array
    wage.gr.bins <- wage.gr.bins.f(SA.base.growth = param.base.growth,
                                   SA.wage.gr = param.wage.gr)
    #data
    df <- get.pop.int()
    #data
    df <- wages.final.cps.org.f(SA.states.raise = param.states.raise, 
                                SA.wages = param.wages, 
                                wage.gr.bins= wage.gr.bins,
                                df.temp = df)
    #table
    table.n.final <- N.final.f(df.temp = df, workers.gr.temp = workers.gr)
    #scalar
    eta.lit <- eta.lit.f(SA.eta.lit = param.eta.lit)
    #scalar
    factor.extrap <- factor.extrap.f(SA.factor.extrap = param.factor.extrap)
    #table
    #stats3 <- final.other.comp()
    stats3 <-  df  %>% 
    filter(pop_of_int == 1) %>% 
    mutate("adult" = ifelse(age>=20, "Adult", "Teen") )  %>% 
    group_by(adult)  %>% 
    summarise("$\\overline{\\%\\Delta w}$" = wtd.mean( ifelse(wages.final <=10.10,
                                              (10.10 - wages.final)/wages.final, NA) , 
                                       na.rm = TRUE, weights = final_weights) * 100 ) 
  
    stats3 <- rbind("$\\eta_{lit}$" = c( eta.lit * factor.extrap , eta.lit ), 
                  "$\\eta_{w \\leq MW'}$" = c( eta.lit * factor.extrap , eta.lit ) / 
                    (table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100),
                  "$F_{adj}$" = c( 4.5, 4.5 ),
                  t(stats3[,-1]) )
  
  aux.1 <- apply(stats3, 2, function(x) x[1] * x[3])
  stats3 <- rbind(stats3, "$\\widetilde{\\eta_{w\\leq MW}}$"= aux.1)
  
  colnames(stats3) <- c("Adult", "Teen")  
  #scalar
    delta.e1 <-  sum( (1 + workers.gr)^3 * 
       table.n.final["Salary workers ($\\hat{ N_{employed} }$) (millions)",1:2] *  param.N *
       table.n.final["% Salary below new MW ($P(\\hat{w} \\leq MW^{1})$)",1:2]/100 * param.fract.minwage *
       ( 1 - table.n.final["% of non compliers ($\\alpha_{1}$)",1:2]/100 - 0) * param.noncomp *
        c( eta.lit.f(param.eta.lit) * factor.extrap.f(param.factor.extrap) , eta.lit.f(param.eta.lit) ) * 
       stats3["$F_{adj}$", "Teen"] * param.F.adj *
       (stats3["$\\overline{\\%\\Delta w}$",1:2]/100) * param.av.wage.var  
    ) + 0.05 * 0.9
      
    abs(delta.e1  -  -0.4776402)/0.4776402
    return(list("delta.e1" = delta.e1, "table.n.final" = table.n.final))
}
# Up to here I am reproducing the same results as above
# results <- SA.part1()
# table.n.final <- results[[2]]
# delta.e1 <- results[[1]]
# ##DONE WITH EMPLOYMENT now use cps asec
# df <- df1


SA.part2 <- function(
  param.wage.gr = 1.0, #anual average wage growth
  param.worker.gr = 1.0, #anual average workers growth
  param.fract.minwage = 1.0, #Fraction of workers earning below the min wage (in ORG only so far)
  param.av.wage.var = 1.0, #Average wage variation for those affected by min wage (only ORG)
  param.wages = 1.0, # Wages of all workers
  param.nonwage.gr = 1.0, 
  param.eta.lit = 1.0, #Elasticity of labor demand for teenagers
  param.ripple = c("scope_below" = 8.7*1.0, "scope_above" = 11.5*1.0, "intensity" = 0.5*1.0),
  param.base.growth = 0.024 * 1.0, #wage growth of lowest decile (before state min wage increase)
  param.factor.extrap = 1.0, #extrapolation factor from teens to adults
  param.noncomp = 1.0, #Rate of non-compliance
  param.F.adj = 1.0, #Adjustment factor for 'appropiate' population
  param.factor.1 = 1.0,
  param.net.benef = 2e9*1.0,
  param.dist.loss = c(0.01, 0.29, 0.70), #c(0.39567218, 0.53851506, 0.06581275)
  param.jobcut = 1){
    # (calls delta.e1 and table.n.final
    # step2: compute policy effects given the effects on employment for each scenario. 
    #variable
    df$pop_of_int <- with(df,
                       (empl == 1 &
                          (selfinc == 0 & selfemp == 0) & 
                          !(incp_wag == 0 | is.na(incp_wag) ) )
                        )
    df <- add.wage.var(df)
    wage.gr <- wage.gr.asec.f(SA.wage.gr = param.wage.gr)
    workers.gr <- workers.gr.asec.f(SA.worker.gr = param.worker.gr)  
    half.gap <- half.gap.asec.f(SA.wage.gr = param.wage.gr, SA.base.growth = param.base.growth)
    wage.gr.bins <- wage.gr.bins.asec.f(SA.wage.gr = param.wage.gr, SA.base.growth = param.base.growth)
    aux.var  <- wtd.quantile(x = df$wage, probs = 1:9/10,weights = df$hhwgt)
    df <- df %>%
        mutate( wage.deciles = cut(wage, c(0, aux.var, Inf) , 
                                right = TRUE, include.lowest = TRUE) ,  
                wage.adj1 =  wage * ( 1 + wage.gr.bins[wage.deciles] )^4) 

#df <- add.wages.1()
    df$wages.final <- with( df, 
                            ifelse(wage.adj1> st.minw.2016[state,] * param.states.raise,
                                   wage.adj1,
                                   st.minw.2016[state,] * param.states.raise) ) * param.wages   
#ripples
  df <-   df %>% 
  mutate( "hhwgt.2016" = hhwgt * (1 + workers.gr)^4 , 
          "below_min" = ifelse(wages.final <= 10.10 & pop_of_int == 1,
                            1, 
                            0),
          "below_min" = ifelse(is.na(below_min),
                            0, 
                            below_min), 
          "new.wage"  = ifelse(wages.final<10.10 & pop_of_int %in% 1, 
                            10.10, 
                            wages.final), 
          "new.wage"  = ifelse(wages.final>10.10 & 
                                 wages.final<param.ripple["scope_above"] & 
                                 pop_of_int==1,
                               wages.final + param.ripple["intensity"] * 
                                 (param.ripple["scope_above"] - wages.final),
                               ifelse(wages.final>param.ripple["scope_below"] & 
                                        wages.final<=10.10 & 
                                        pop_of_int==1,
                                      10.10 + param.ripple["intensity"] * 
                                        (wages.final - 7.25),
                                      new.wage
                          ) )
  )
    alpha.1 <- table.n.final["% of non compliers ($\\alpha_{1}$)", "Total"] * param.noncomp /100
    df <- add.nocomp(df.temp = df, alpha.1.temp = alpha.1)
#job.kill function
  set.seed(123)
  job.cut.factor <- 2 * param.jobcut
  num.to.del <- - delta.e1*1e6*job.cut.factor
  #WOULD LIKE TO MAKE THE NEXT OPTIMIZATION MORE EFFICIENT
  theta <- 3
  cut.job <- 0
  while ( abs(sum(cut.job * df$hhwgt.2016 , na.rm = TRUE) - num.to.del) >= 1e5)  { 
    set.seed(123)
    cut.job <- 1* ( df$wages.final<=10.10 & df$pop_of_int==1 & df$no.comply=="comply" &
                      (theta*runif(dim(df)[1], min = 0, max = 1) < 0.5) )
    # sum(cut.job * df$hhwgt.2016 , na.rm = TRUE)/1e6
    if (sum(cut.job * df$hhwgt.2016 , na.rm = TRUE) - num.to.del >= 10000) {
      theta <- theta * 1.001 
    } else {
      theta <- theta * 0.999
    }
     #print(theta)
  } 
  df$cut.job <- cut.job
  df$cut.job[is.na(df$cut.job)] <- 0
  rm(cut.job)
  
  #df$teen <- df$age<20
  #prop.table(with(df[df$pop_of_int == 1, ], table(cut.job,teen)), 2)
  
  # Compute variables for each scenario
  # cut jobs, no wage raise
  df$cut.wage <- with(df, ifelse(cut.job %in% 1, 
                                 wages.final/job.cut.factor, 
                                 wages.final) )
  # cut jobs relative to sq, wage raise
  df$new.wage.final <- with(df, ifelse(cut.job %in% 1, 
                                       wages.final/job.cut.factor, 
                                       new.wage.nocomp) )
  # cut jobs relative to raise, wage raise
  df$new.wage.cut <- with(df, ifelse(cut.job %in% 1, 
                                     new.wage.nocomp/job.cut.factor, 
                                     new.wage.nocomp) )
    
  non.wage.gr <- non.wage.gr.f(SA.nonwage.gr = param.nonwage.gr)

                df <- all.income.f(df.temp = df, non.wage.gr.temp = non.wage.gr)
}  

# Compute variation by hhld - plot all the effects in same units (average per group) with quintiles instead of poverty lines.
final.plot <- function(df.temp = df) {
      final_fig3 <- df.temp %>% 
      select(new.inc.pc,sq.inc.pc, hhwgt.2016, balance.loss, winners, losers, income.group)  %>%   
      mutate( "variation" = new.inc.pc - sq.inc.pc, 
               "inc_gain" = ifelse(variation>0, 
                                  "gain", 
                                  ifelse(variation < 0, "loss", "same" ) ) ) %>% group_by(income.group) %>% 
      summarise("mean" = wtd.mean(variation, weights = hhwgt.2016),
                "mean (win)" = wtd.mean(winners, weights = hhwgt.2016), 
                "mean (lose)" = wtd.mean(losers, weights = hhwgt.2016),
                "mean (bal lose)" = wtd.mean(balance.loss, weights = hhwgt.2016),
                "N" = sum( hhwgt.2016)   
                ) 
    
    high.loss <-  as.numeric(final_fig3[max(final_fig3$income.group), "mean (bal lose)"])
    
    final_fig3 <- final_fig3 %>% 
      select(`mean (win)`, `mean (lose)`, `mean (bal lose)`, income.group) %>% 
      melt(value.variables=c("mean (win)", "mean (lose)") , id="income.group") %>% 
      ggplot(aes(as.factor(income.group), value, fill = as.factor(variable))) + 
      geom_bar(color = "gray", alpha = 0.5, 
               stat = "summary", fun.y = "mean", 
               position = "dodge") +
      coord_cartesian(ylim = c(0,400) ) +
      geom_text(x = 4.8, y = 365,angle = 90, 
                label = - round(high.loss) , 
                size = 3, colour = "#6699FF", alpha = 0.5) + 
      geom_segment(aes(x = 5, y = 300, xend = 5, yend = 410), 
                   colour = "#6699FF", arrow = arrow(length = unit(0.2, "cm"))) + 
      theme(legend.justification=c(0, 0), 
                    legend.position=c(0, .7),
                    legend.background = element_rect(colour = 'transparent', fill = 'transparent') 
                    ) +
      scale_fill_discrete(name=NULL, 
                          labels=c("Wage +", "Wage -", "Balance -")) + 
      labs(y = "$/year", 
           x = "Quintiles of per capita income" , 
           title = "Distribution of gains and losses across quintiles") +
  theme(plot.title = element_text(hjust = 0.5))   
    
    #two outputs of the SA
    return(final_fig3)
}

###########################
###### HERE STARTS THE SA: MODIFY CORRESPONDING PARAMETERS TO REPRODUCE TABLE 1 FROM PAPER
df <- df_csporg
results <- SA.part1()
table.n.final <- results[[2]]
delta.e1 <- results[[1]]
##DONE WITH EMPLOYMENT now use cps asec
df <- df_asec
df3 <- SA.part2()

param.factor.1 <- 1.0
param.net.benef <- 2e9*1.0
#param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.39567218, 0.53851506, 0.06581275)
param.dist.loss <- c(0.39567218, 0.53851506, 0.06581275)


losses <- with(df3, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) ) - param.net.benef
pop.dist <- wtd.table( with(df3, 
                            findInterval(x = sq.inc.pc, 
                                         vec = c(-Inf,11740, 6*11740, Inf)) ) ,
                      weights = df3$hhwgt.2016 )$sum.of.weights
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc, 
                       df.temp = df3)
    
results2 <- (sapply( c(-0.1, 0.1), 
                  function(y) final_dec.f(x = y, df.temp = df.final) ) -v0 )/v0
results2
df4 <- df.final
quintiles <- with(df4,wtd.quantile(x = sq.inc.pc, probs = 1:4/5, weights = hhwgt.2016))
df4$income.group <- with(df4, findInterval(x = sq.inc.pc, vec =  c(-Inf,quintiles) ))
p <- final.plot(df.temp = df4)
print(p)

#    print("Percentage variation relative to base scenario:")
#    print( (sapply(c(-0.1, 0.1), function(y) final_dec.f(df.temp = df4, y) ) - v0)/v0 )

```  



```{r SA f.extrap, eval=FALSE, echo=FALSE} 
#param.factor.extrap = 3 * seq(0,1.5, length.out = 10)
moving.p <- 3 * 1/3
grid.num <- 20
welf.vals <- matrix(nrow = grid.num, ncol = 2, data = NA)
colnames(welf.vals) <- c("rho_neg", "rho_pos")
factor.extrap.vals <- seq(0,1.5, length.out = grid.num)

i <- 0                       
for (moving.p in 3 * factor.extrap.vals) {
i <- i + 1
df <- df_csporg
results <- SA.part1(param.factor.extrap = moving.p)
table.n.final <- results[[2]]
delta.e1 <- results[[1]]
##DONE WITH EMPLOYMENT now use cps asec
df <- df_asec
df3 <- SA.part2(param.factor.extrap = moving.p)

param.factor.1 <- 1
param.net.benef <- 2e9*1.0
param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.39567218, 0.53851506, 0.06581275)

losses <- with(df3, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) ) - param.net.benef
pop.dist <- wtd.table( with(df3, 
                            findInterval(x = sq.inc.pc, 
                                         vec = c(-Inf,11740, 6*11740, Inf)) ) ,
                      weights = df3$hhwgt.2016 )$sum.of.weights
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc, 
                       df.temp = df3)
welf.vals[i,] <- sapply( c(-0.1, 0.1), 
                  function(y) final_dec.f(x = y, df.temp = df.final) ) 
}

df.plot1 <- tbl_df(data.frame(cbind(factor.extrap.vals, welf.vals)))

final_fig1 <- df.plot1 %>%
  ggplot(aes(factor.extrap.vals, rho_neg)) + 
  coord_cartesian(xlim = c(-0.1,1.6), 
                    ylim = c(-13,13),expand = FALSE ) +
  scale_x_discrete(limits = c(0, 1/3, 1.5), 
                   labels = c("0", "1/3", "1.5")) + 
  scale_y_discrete(limits = c(-12, 0, 12), 
                   labels = c("-12", "0", "12")) +
  geom_hline(color = "black", yintercept = 0, size = 0.5, linetype="dotted") +
  geom_line( colour = "black", size = 0.9, linetype = "dotdash") + 
  geom_line(aes(factor.extrap.vals, rho_pos), colour = "red", size = 0.7) + 
  geom_text(x = 1.2, y = -4,angle = 0, 
                label = "rho = -0.1 (dislikes)", 
                size = 3, colour = "black") + 
  geom_text(x = 1.2, y = 8,angle = 0, 
              label = "rho = 0.1 (likes)", 
              size = 3, colour = "red") + 
  labs(y = "W", 
       x = "x: elas_adult = x * elas_teen" , 
       title = "Extrapolation Factor") + 
  theme(plot.title = element_text(hjust = 0.5) )
print(final_fig1)
ggsave("sa_extrap.png")
```

```{r SA F.adj, eval=FALSE, echo=FALSE} 
#param.factor.extrap = 3 * seq(0,1.5, length.out = 10)
moving.p <- 1/4.5 
grid.num <- 20
welf.vals <- matrix(nrow = grid.num, ncol = 2, data = NA)
colnames(welf.vals) <- c("rho_neg", "rho_pos")
factor.adj.vals <- seq(1,20, length.out = grid.num)

i <- 0                       
for (moving.p in 1/4.5 * factor.adj.vals) {
i <- i + 1
df <- df_csporg
results <- SA.part1(param.F.adj = moving.p)
table.n.final <- results[[2]]
delta.e1 <- results[[1]]
##DONE WITH EMPLOYMENT now use cps asec
df <- df_asec
df3 <- SA.part2(param.F.adj = moving.p)

param.factor.1 <- 1
param.net.benef <- 2e9*1.0
param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.39567218, 0.53851506, 0.06581275)

losses <- with(df3, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) ) - param.net.benef
pop.dist <- wtd.table( with(df3, 
                            findInterval(x = sq.inc.pc, 
                                         vec = c(-Inf,11740, 6*11740, Inf)) ) ,
                      weights = df3$hhwgt.2016 )$sum.of.weights
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc, 
                       df.temp = df3)
welf.vals[i,] <- sapply( c(-0.1, 0.1), 
                  function(y) final_dec.f(x = y, df.temp = df.final) ) 
}

df.plot2 <- tbl_df(data.frame(cbind(factor.adj.vals, welf.vals)))

final_fig1 <- df.plot2 %>%
  ggplot(aes(factor.adj.vals, rho_neg)) + 
  coord_cartesian(xlim = c(-0.1,21), 
                    ylim = c(-13,13),expand = FALSE ) +
  scale_x_discrete(limits = c(0, 4.5, 20), 
                   labels = c("0", "4.5", "20")) + 
  scale_y_discrete(limits = c(-12, 0, 12), 
                   labels = c("-12", "0", "12")) +
  geom_hline(color = "black", yintercept = 0, size = 1/3, linetype="dotted") +
  geom_line( colour = "black", size = 0.9, linetype = "dotdash") + 
  geom_line(aes(factor.adj.vals, rho_pos), colour = "red", size = 0.7) + 
  labs(y = "W", 
       x = "x: elas_lit = F_adjs * elas_teen" , 
       title = "Adjustment Factor") +
  theme(plot.title = element_text(hjust = 0.5))
print(final_fig1)
ggsave("sa_f_adj.png")
#Saving 7.28 x 4.26 in image

```

```{r SA eta.teen, eval=FALSE, echo=FALSE} 
#param.factor.extrap = 3 * seq(0,1.5, length.out = 10)
moving.p <- 1/(0.10) 
grid.num <- 10
welf.vals <- matrix(nrow = grid.num, ncol = 2, data = NA)
colnames(welf.vals) <- c("rho_neg", "rho_pos")
factor.eta.vals <- seq(0,0.20, length.out = grid.num)

i <- 0                       
for (moving.p in 1/(0.10) * factor.eta.vals) {
i <- i + 1
df <- df_csporg
results <- SA.part1(param.eta.lit = moving.p)
table.n.final <- results[[2]]
delta.e1 <- results[[1]]
##DONE WITH EMPLOYMENT now use cps asec
df <- df_asec
df3 <- SA.part2(param.eta.lit = moving.p)

param.factor.1 <- 1
param.net.benef <- 2e9*1.0
param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.39567218, 0.53851506, 0.06581275)

losses <- with(df3, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) ) - param.net.benef
pop.dist <- wtd.table( with(df3, 
                            findInterval(x = sq.inc.pc, 
                                         vec = c(-Inf,11740, 6*11740, Inf)) ) ,
                      weights = df3$hhwgt.2016 )$sum.of.weights
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc, 
                       df.temp = df3)
welf.vals[i,] <- sapply( c(-0.1, 0.1), 
                  function(y) final_dec.f(x = y, df.temp = df.final) ) 
}

df.plot3 <- tbl_df(data.frame(cbind(factor.eta.vals, welf.vals)))

final_fig1 <- df.plot3 %>%
  ggplot(aes(factor.eta.vals, rho_neg)) + 
  coord_cartesian(xlim = c(0,0.21), 
                    ylim = c(-13,13),expand = FALSE ) +
  scale_x_discrete(limits = c(0, 0.10, 0.20), 
                   labels = c("0", "0.10", "0.20")) + 
  scale_y_discrete(limits = c(-12, 0, 12), 
                   labels = c("-12", "0", "12")) +
  geom_hline(color = "black", yintercept = 0, size = 1/3, linetype="dotted") +
  geom_line( colour = "black", size = 0.9, linetype = "dotdash") + 
  geom_line(aes(factor.eta.vals, rho_pos), colour = "red", size = 0.7) + 
  labs(y = "W", 
       x = "elas_teen" , 
       title = "Elasticy of Labor Demand for Teenagers") +
  theme(plot.title = element_text(hjust = 0.5))
print(final_fig1)
ggsave("sa_eta.png")
```

```{r SA dist.loss, eval=FALSE, echo=FALSE} 
df <- df_csporg
results <- SA.part1(param.F.adj = 1)
table.n.final <- results[[2]]
delta.e1 <- results[[1]]
##DONE WITH EMPLOYMENT now use cps asec
df <- df_asec
df3 <- SA.part2()

param.factor.1 <- 1
param.net.benef <- 2e9*1.0
param.dist.loss <- c(0.01, 0.29, 0.70) #c(0.39567218, 0.53851506, 0.06581275)

losses <- with(df3, sum(winners * hhwgt.2016) - 
                 param.factor.1 * sum(losers * hhwgt.2016) ) - param.net.benef
pop.dist <- wtd.table( with(df3, 
                            findInterval(x = sq.inc.pc, 
                                         vec = c(-Inf,11740, 6*11740, Inf)) ) ,
                      weights = df3$hhwgt.2016 )$sum.of.weights
losses.pc  <- as.numeric(losses) * param.dist.loss / pop.dist

df.final <- win.loss.f(losses.pc.temp = losses.pc, 
                       df.temp = df3)
w1 <- final_dec.f(0.1, df.temp = df.final)  
w2 <- final_dec.f(-0.1, df.temp = df.final)  

#dev.off()
png("sa_DLf.png", width = 4, height = 4, units = 'in', res = 600)
par(oma=c(2,2,3,0) + 0.1,mar=c(0,0,0,0) + 0.1)
layout(matrix(c(1,1,1,1,2,3,4,5), 4, 2, byrow = FALSE), widths=c(1,3), heights=c(1,1,1,1))
barplot(losses.pc, pop.dist, space=0, 
        col = rgb(0,0,0,.3), xaxt='n' )
segments(x0 = 2e8, x1 = 10e8, y0 = 200, y1 = 200,  col = "red", lwd = 2, lty = 3)
text(x = 1.5e8, y = 800, labels =  paste(round(w1),"/",round(w2))  , cex = 1.4)

losses.pc1  <- as.numeric(losses) * c(0.01, 0.05, 0.94) / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc1, 
                       df.temp = df3)

w1 <- final_dec.f(0.1, df.temp = df.final)  
w2 <- final_dec.f(-0.1, df.temp = df.final)  
barplot(losses.pc1, pop.dist, space=0, 
        col = rgb(0,0,0,.3), ylim = c(0,200), xaxt='n', yaxt='n' )
segments(x0 = 2.6e8, x1 = 10e8, y0 = 200, y1 = 200,  col = "red", lwd = 2, lty = 3)
text(x = 1.9e8, y = 130, labels =  paste(round(w1),"/",round(w2))  , cex = 1.4)

losses.pc2  <- as.numeric(losses) * c(0.30, 0.50, 0.20) / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc2, 
                       df.temp = df3)

w1 <- final_dec.f(0.1, df.temp = df.final)  
w2 <- final_dec.f(-0.1, df.temp = df.final)  
barplot(losses.pc2, pop.dist, space=0, 
        col = rgb(0,0,0,.3), ylim = c(0,200), xaxt='n', yaxt='n' )
segments(x0 = 3e8, x1 = 3.3e8, y0 = 200, y1 = 200,  col = "red", lwd = 2, lty = 3)
text(x = 1.9e8, y = 130, labels =  paste(round(w1),"/",round(w2))  , cex = 1.4)

losses.pc3  <- as.numeric(losses) * c(0.39567218, 0.53851506, 0.06581275) / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc3, 
                       df.temp = df3)
w1 <- final_dec.f(0.1, df.temp = df.final)  
w2 <- final_dec.f(-0.1, df.temp = df.final)  
barplot(losses.pc3, pop.dist, space=0, 
        col = rgb(0,0,0,.3), ylim = c(0,200), yaxt='n' , xaxt = 'n')
segments(x0 = 3e8, x1 = 3.3e8, y0 = 200, y1 = 200,  col = "red", lwd = 2, lty = 3)
text(x = 1.9e8, y = 130, labels =  paste(round(w1),"/",round(w2))  , cex = 1.4)

losses.pc4  <- as.numeric(losses) * c(.45, .50, 0.05) / pop.dist
df.final <- win.loss.f(losses.pc.temp = losses.pc4, 
                       df.temp = df3)
w1 <- final_dec.f(0.1, df.temp = df.final)  
w2 <- final_dec.f(-0.1, df.temp = df.final)  
barplot(losses.pc4, pop.dist, space=0, 
          col = rgb(0,0,0,.3), ylim = c(0,200), yaxt='n' , xaxt ='n')
axis(side = 1, at = c(0.5e8, 1.9e8, 3.1e8), labels = c("<1PL", "1PL,6PL", ">6PL") ) 
segments(x0 = 3e8, x1 = 3.3e8, y0 = 200, y1 = 200,  col = "red", lwd = 2, lty = 3)
text(x = 1.9e8, y = 130, labels = paste(round(w1),"/",round(w2)) , cex = 1.4)
mtext("Distribution of Balance Losses and Welfare \n W(0.1) / W(-0.1)", outer = TRUE, cex = 1)
dev.off()
```


[^1]: The report presents smaller estimates for the \$9.00 dollar option (-0.075). The rationale is that a smaller increase (in magnitude but also not indexed by inflation, and with later implementation than the 10.10 option)   will allow firms to adjust other margins before reducing employment. 

[^2]: CBO calculated the fraction of teenagers with earnings below the minimum wage from 1979 to 2009 and the result came to about a third. Then they look at the average change in earnings for teenagers subject to the minimum wage over the same period, and compared that to the nominal change in each variation of the minimum wage. This ratio came to be about 1.5. With this the final estimates for the elasticity for teenagers came to be 4.5 ($1.5/(1/3)$) times higher than what is estimated in the literature.   

